---
title: "Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics"
affiliation:
  author-columnar: true         ## one column per author
  #institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields
  institution:
    - name: Queensland University of Technology
      department: Science and Engineering Faculty
      location: Brisbane, Australia
      email: stephanie.kobakian@gmail.com
      mark: 1
      author:
        - name: Stephanie Kobakian
    - name: Monash University
      department: Econometrics and Business Statistics Faculty
      location: Melbourne, Australia
      email: dicook@monash.edu
      mark: 1
      author:
        - name: Dianne Cook
keywords: ["statistics", "visual inference", "geospatial", "population"]
abstract: |
  The choropleth map display is commonly used for communicating spatial distributions across geographic areas. However, when choropleths are used the size of the geographic units will influence the understanding of the distribution derived by map users. The hexagon tile map is presented as an alternative display for visualizing population related distributions effectively. Visual inference is used to measure the power of the hexagon tile map design, and the choropleth is used as a comparison. The hexagon tile map display is tested using a distribution that is directly related to the geography, with values monotonically increasing from the North-West to South-East areas of Australia. This study finds in a hexagon tile map lineup the single map that contains a population related distribution is detected with greater probability than the same data displayed in a choropleth map. These findings should encourage map creators to implement alternative displays and consider a hexagon tile map when presenting spatial distributions of heterogeneous areas.
bibliography: paper.bib
output:
  pdf_document: 
      number_sections: true
      keep_tex: true
editor_options: 
  chunk_output_type: console
link-citations: true
citation_package: biblatex
csl: springer-vancouver.csl
header-includes: \usepackage{graphicx}
---

<!-- Regular paper – 12 double column pages (Submissions may be up to 18 pages in length, subject to MOPC. All regular paper page limits include references and author biographies.) -->

<!-- Changes: thyroid instead of liver - check; lineup need to be hexmap; check language with chatgpt/grammarly; get refs formatted and paper into 10 pages; use new template -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  error = FALSE, 
  message = FALSE,
  cache = FALSE,
  dpi = 300,
  out.width = "100%")
```

```{r libraries}
# Load Libraries
library(tidyverse)
library(readxl)
library(broom)
library(broom.mixed)
library(cowplot)
library(png)
library(grid)
library(lme4)
library(ggthemes)
library(RColorBrewer)
library(knitr)
library(kableExtra)

options(knitr.kable.digits = "2")
```

```{r data}
trend_colors <- c(
  "NW-SE" = "#B2DF8A",
  "Three Cities" = "#A6CEE3",
  "All Cities" = "#1F78B4")
  
type_colors <- c(
  "Choro." = "#fcae91",
  "Hex." = "#a50f15")

detect_f_colors <- c(
  "No" = "#66C2A5",
  "Yes" = "#FC8D62")

detect_colors <- c(
  "Detected? No" = "#66C2A5",
  "Detected? Yes" = "#FC8D62")

  # Downloaded data
d <- read_csv("data/experiment-data.csv")

# Check data set 
# Need to clean multiple entries, 48, 24
# remove duplicated entries due to submit button
d <- d %>% group_by(group, pid, image_name) %>%
  slice(1) %>% ungroup() %>% 
  arrange(group, pid, plot_order)

# Remove pids who did not provide answers to most questions
keep <- d %>% count(pid, sort = TRUE) %>% filter(n > 10)
d <- d %>% 
  filter(pid %in% keep$pid) %>%
  filter(pid != "1234567890")

# Remove pids who did not provide any choices
# Or an insufficient amount of responses
bad_contribs <- d %>% group_by(pid) %>% 
  summarise(sum0 = sum(choice)) %>% 
  filter(sum0 < 13) %>% 
  pull(pid)

d <- d %>% 
  filter(!(pid %in% bad_contribs))


n_pids <- d %>% count(pid, sort=TRUE) %>% 
  summarise(n_pids = length(pid))

d <- d %>% mutate(certainty = factor(as.character(certainty),
  levels = c("1", "2", "3", "4","5"), ordered=TRUE))
```


```{r reps}
replicate <- tibble(image_name = c("aus_cities_12_geo.png", "aus_cities_12_hex.png", 
                                   "aus_cities_3_geo.png", "aus_cities_3_hex.png",
                                   "aus_cities_4_geo.png", "aus_cities_4_hex.png",
                                   "aus_cities_9_geo.png", "aus_cities_9_hex.png",
                                   "aus_nwse_2_geo.png", "aus_nwse_2_hex.png",
                                   "aus_nwse_3_geo.png", "aus_nwse_3_hex.png",
                                   "aus_nwse_5_geo.png", "aus_nwse_5_hex.png",
                                   "aus_nwse_6_geo.png", "aus_nwse_6_hex.png",
                                   "aus_three_12_geo.png", "aus_three_12_hex.png",
                                   "aus_three_5_geo.png", "aus_three_5_hex.png",
                                   "aus_three_8_geo.png", "aus_three_8_hex.png",
                                   "aus_three_9_geo.png", "aus_three_9_hex.png"),
                    replicate = c(1, 1, 2, 2, 3, 3, 4, 4, 
                                  1, 1, 2, 2, 3, 3, 4, 4,
                                  1, 1, 2, 2, 3, 3, 4, 4))
# Add rep info to data
d <- d %>% left_join(., replicate, by = "image_name")
```

```{r pdetection_group}
# Tidy for analysis
d <- d %>% 
  separate(image_name, c("nothing", "trend", "location", "type", "extra"), remove = FALSE) %>%
  select(-nothing, -extra) %>%
  mutate(location = as.numeric(location), 
    # detect measures the accuracy of the choice
         detect = ifelse(location == choice, 1, 0)) %>% 
  mutate(trend = case_when(
    trend == "nwse" ~ "NW-SE",
    trend == "cities" ~ "All Cities",
    trend == "three" ~ "Three Cities")) %>% 
  mutate(trend = fct_relevel(trend,"All Cities","Three Cities", "NW-SE")) %>% 
  mutate(type = case_when(
    type == "hex" ~"Hex.",
    TRUE~"Choro.")) %>% 
    mutate(detect_f = factor(detect, levels = c(0,1), labels = c("Detected? No", "Detected? Yes")))

plots <- d %>% group_by(group, trend, type, location) %>%
  # pdetect measures the aggregated accuracy of the choices
  summarise(pdetect = length(detect[detect == 1])/length(detect)) 
```

```{r makethyroidplots}
library(cartogram)
library(sugarbag)
library(sf)

# Spatial polygons files sourced from: https://github.com/wfmackey/absmapsdata
load("data/sa22011.rda")
load("data/state2011.rda")

invthm <- theme_map() + 
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA), 
    plot.background = element_rect(fill = "black", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.key = element_rect(fill = "transparent", colour = NA),
    text = element_text(colour = "white"),
    axis.text = element_blank()
  )

# function to allocate colours to regions
aus_colours <- function(sir_p50){
  value <- case_when(
    sir_p50 <  0.74 ~ "#33809d",
    sir_p50 >= 0.74 & sir_p50 < 0.98 ~ "#aec6c7",
    sir_p50 >= 0.98 & sir_p50 < 1.05 ~ "#fff4bc",
    sir_p50 >= 1.05 & sir_p50 < 1.45 ~ "#ff9a64",
    sir_p50 >= 1.45 ~ "#ff3500",
    TRUE ~ "#FFFFFF")
  return(value)
}

sa2 <- sa22011 %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(!state_name_2011 == "Other Territories") %>% 
  filter(!sa2_name_2011 == "Lord Howe Island")

SIR <- read_csv("data/SIR Downloadable Data.csv") %>% 
  filter(SA2_name %in% sa2$sa2_name_2011) %>% 
  dplyr::select(Cancer_name, SA2_name, Sex_name, p50) %>% 
  filter(Cancer_name == "Thyroid", Sex_name == "Females")
ERP <- read_csv("data/ERP.csv") %>%
  filter(REGIONTYPE == "SA2", Time == 2011, Region %in% SIR$SA2_name) %>% 
  dplyr::select(Region, Value)
# Alternative maps
# Join with sa2 sf object
sa2thyroid_ERP <- SIR %>% 
  left_join(sa2, ., by = c("sa2_name_2011" = "SA2_name")) %>%
  left_join(., ERP %>% 
              dplyr::select(Region, 
              Population = Value), by = c("sa2_name_2011"= "Region")) %>% 
  filter(!st_is_empty(geometry))
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  #filter(!is.na(Population)) %>% 
  filter(!sa2_name_2011 == "Lord Howe Island") %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf() 

# Make choropleth
aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR), size = 0.1) + 
  scale_fill_identity() + invthm

# Make hexmap
if (!file.exists("data/aus_hexmap.rda")) {

## Check the projection uses long and lat
## Create centroids set
sa2centroids <- sa2 %>% 
  create_centroids(., "sa2_name_2011")
## Create hexagon grid
# For speed, consider if the buffer distance beyond the centroids is necessary
grid <- create_grid(centroids = sa2centroids,
                    hex_size = 0.15,
                    buffer_dist = 4, verbose = TRUE)
## Allocate polygon centroids to hexagon grid points
# This is a time consuming process for a large data set
aus_hexmap <- allocate(
  centroids = sa2centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.15,
  hex_filter = 8,
  focal_points = capital_cities,
  width = 25,
  verbose = TRUE)
save(aus_hexmap, file = "data/aus_hexmap.rda") 
}

load("data/aus_hexmap.rda")

## Prepare to plot
fort_hex <- fortify_hexagon(data = aus_hexmap,
                            sf_id = "sa2_name_2011",
                            hex_size = 0.3) %>% 
            left_join(sa2thyroid_ERP %>%
            select(sa2_name_2011, SIR, p50))
fort_aus <- sa2thyroid_ERP %>%
  fortify_sfc()
## Make a plot
aus_hexmap_plot <- ggplot() +
  geom_polygon(aes(x = long,  y = lat,  
        group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey", 
               linewidth = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, 
                   fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()
```

Introduction
=============
<!-- General: comparison of displays, motivate hexagon tile map, enough info to motivate aims -->
<!-- geographies on choropleth -->
This study compares the effectiveness of the spatial display, a hexagon tile map, against the standard, a choropleth map, for communicating information about disease statistics. The choropleth map is the traditional method for visualizing aggregated statistics across administrative boundaries. The hexagon tile map builds on existing displays, such as the cartogram, and tessellated hexagon displays. A hexagon tile map forgoes the familiar boundaries, in favour of representing each geographic unit as an equally sized hexagon, placed approximately in the correct spatial location.  It differs in the relaxed requirement to have connected hexagons, and allows sparsely located hexagons. This type of display may be useful for other countries, and other purposes. The algorithm to construct a hexagon tile map is available in the R package sugarbag [@sugarbag]. 

The hexagon tile map was designed for Australia, motivated by a need to display spatial statistics for the Australian Cancer Atlas. None of the existing approaches for creating cartograms or hexagon tiling perform well for the Australian landscape, which has vast open spaces and concentrations of population in small regions clustered on the coastlines. XXX cite Kobakian, Cook & Roberts 2020 and explain approach to reach this conclusion 

The Australian Cancer Atlas [@atlas] is an online interactive web tool created to explore the burden of cancer on Australian communities. There are many cancer types to be explored individually or aggregated. XXX Who are the users of the Atlas

The Australian Cancer Atlas allows users to explore the patterns in the distributions of cancer statistics over the geographic space of Australia. It uses a choropleth map display and diverging colour scheme to draw attention to relationships between neighbouring areas. The hexagon tile map may be a useful alternative display to enhance the atlas. 

The experiment was conducted using the lineup protocol, a visual inference procedure [@GIIV], to objectively test the effectiveness of the two displays. XXX explain lineup protocol with more detail

The paper is organised as follows. The next section discusses the background of geographic data display and visual inference procedures. The [Methodology] section describes the methods for conducting the experiment and analysing the results. The results are summarized in the [Results] section, followed by a discussion about the broader implications for the use of this map style.


Background
=============

## Spatial data displays

<!-- foundation, tradition -->  
Spatial visualisations communicate the distribution of statistics over geographic landscapes. The choropleth map [@EI], [@BCM] is a traditional display. It is used to present statistics that have been aggregated on geographic units. Creating a choropleth map involves drawing polygons representing the administrative boundaries, and filling with colour mapped to the value of the statistic. The choropleth map places the statistic in the context of the spatial domain, so that the reader can see whether there are spatial trends, clusters or anomalies. 

XXX make it clear that visualizing communicable disease patterns is just one example use of a spatial data display.
This is important for digesting disease patterns. If there is a trend it may imply that the disease is spreading from one location to another. If there is a cluster, or an anomaly, there may be a localized outbreak of the disease. Aggregating the statistic on administrative units, provides a level of privacy to individuals, while allowing the impact of the disease on the community to be analysed.

<!-- Many sets of Australian statistical areas have a large disparity between the smallest and largest units.
The rural communities in Australia operate on a much larger geographic space than small inner-city communities. When a spatial distribution is related to the size of the areas, or the population density, the size of the regions can allow erroneous conclusions to be drawn about the state of the statistic over the entire population, when presenting the statistics on a geographic map base. This occurs as large regions filled with a consistent colour or pattern can draw the attention of map readers, and small regions are not paid equal attention [@CTTMB]. An alternative display can be used to honestly communicate a spatial distribution for a set of heterogeneous areas [@NISCC]. 
This work aims to show that a hexagon tile map display is a viable alternative to the geographic map base for presenting population statistics.


A choropleth map is not the only display that can be used for presenting geospatial data. 
Alternative maps include various cartograms and tessellated tile maps. They allow other variables to be incorporated in the display to highlight the statistical values of various geographic areas.
-->

```{r liver, eval=FALSE, fig.height = 4, fig.width = 4, fig.cap = "A choropleth map of the smoothed average of liver cancer diagnoses for Australian males. The diverging colour scheme uses dark blue areas for much lower than average diagnoses, yellow areas with diagnoses around the Australian average, red shows diagnoses much higher than average. The hexagon tile map shows concentrations of higher than expected liver cancer rates in the cities of Melbourne and Sydney, which is not visible from the choropleth.", results = "asis"}
ggdraw() +
  draw_plot(rasterGrob(readPNG("figuresaus_liver_m.png")))
```

```{r liver-hex, eval=FALSE, fig.height = 4, fig.width = 4, fig.cap = "A hexagon tile map of the smoothed average of liver cancer diagnoses for Australian males. The diverging colour scheme uses dark blue areas for much lower than average diagnoses, yellow areas with diagnoses around the Australian average, red shows diagnoses much higher than average. The hexagon tile map shows concentrations of higher than expected liver cancer rates in the cities of Melbourne and Sydney, which is not visible from the choropleth.", results = "asis"}

ggdraw() +
  draw_plot(rasterGrob(readPNG("figuresaus_liver_m_hex.png")))
```
<!--
This will create the figure as needed in the paper, but I can't get the width argument to work properly.
\begin{figure*}[!t]
\centering
\includegraphics{paper_files/figure-latex/liver-1.pdf}
\caption{A choropleth map (a), and hexagon tile map (b) of the smoothed average of liver cancer diagnoses for Australian males. The diverging colour scheme uses dark blue areas for much lower than average diagnoses, yellow areas with diagnoses around the Australian average, red shows diagnoses much higher than average. The hexagon tile map shows concentrations of higher than expected liver cancer rates in the cities of Melbourne and Sydney, which is not visible from the choropleth.}
\label{fig:liver2}
\end{figure*}
-->

```{r thyroid-choro, out.height = "30%", fig.height = 3, fig.width = 4, fig.align='center', fig.cap = "A choropleth map of thyroid cancer incidence among females across the Statistical Areas of Australia at Level 2. Blue indicates lower than average and red indicates higher than average incidence. A cluster of high incidence is visible on the east coast."}
aus_ggchoro
```

The choropleth map is an effective spatial display if the size of the geographic units is relatively uniform. This is not the case for most countries. Size heterogeneity in administrative units is particularly extreme in Australia: most of the landscape of Australia is sparsely settled, with the population densely clustered into the narrow coastal strips. Fig. \ref{fig:thyroid-choro} shows the choropleth map of thyroid cancer rates in Australia. The choropleth map focuses attention on the geography, and for heterogeneously sized areas it presents a biased view of the population related distribution of the statistic [@CBATCC]. *Land doesn't get cancer, people do* -- a more effective way to communicate the spatial distributions of cancer statistics is needed. XXX add reference

<!-- Cartogram -->
A cartogram is a general solution for adequately displaying a population-based statistic. It transforms the geographic map base to reflect the population in the geographic region, while preserving some aspects of the geographic location. There are several cartogram algorithms [@ACTUC], [@CBATCC]; each involves shifting the boundaries of geographic units, using the value of the statistic to increase or decrease the area taken by the geographic unit on the map. The changes to the boundaries result in cartograms that accurately communicate population by map area for each of the geographic units but can result in losing the familiar geographic information. For Australia, the transformations warp the country so that it is no longer recognisable.

<!-- Other cartograms -->
Alternative algorithms make various trade offs between familiar shapes and representation of geographic units. The non-contiguous cartogram method [@NAC] keeps the shapes of geographic units intact, and changes the size of the shape. This method disconnects areas creating empty space on the display losing the continuity of the spatial display of the statistic. The Dorling cartogram [@ACTUC] represents each unit as a circle, sized according to the value of the statistic. The neighbour relationships are mostly maintained by how the circles touch. A similar approach was pioneered by Raisz [-@RSCW], using rectangles that tile to align borders of neighbours [@CDWCS]. There have been thorough reviews of the array of methods, as suitable for cancer atlas displays [@review], [@BCM].

```{r thyroid-hex, out.height = "30%", fig.height = 3, fig.width = 4, fig.align='center', fig.cap = "A hexagon tile map of female thyroid cancer incidence in Australia, the same data as shown in the choropleth map in Figure 1. The high incidence in several of the metropolitan regions (Brisbane, Sydney and Perth) can now be seen, along with numerous isolated spots."}
aus_hexmap_plot
```

<!-- Hexagon tile map -->
The hexagon tile map algorithm, automatically matches spatial regions to their nearest hexagon tile, from a grid of tiles. It has the effect of spreading out the inner city areas while maintaining the spatial locations or regions in remote areas. The algorithm is available in the R package, sugarbag [@sugarbag]. Fig. \ref{fig:thyroid-hex} shows the hexagon tile map where the map is coloured according to the thyroid cancer incidence rate in the area. The colour scheme used is bidirectional, it ranges from substantially below average (blue) to substantially above average (red) for the cancer incidence rate within the geographic area. In the hexagon tile map display, the inner city areas have expanded, making it possible to see the cancer incidence in the small, densely populated areas. Remote regions are represented by isolated hexagons, which is not ideal, but maintains the spatial location of these data values and highlights the sparsity of residents. It is of interest to know how well the spatial distribution is perceived by map users for this display, in comparison to the choropleth.
XXX clarify who is doing the perceiving 

## Visual Inference

In order to assess the effectiveness of the hexagon tile map, the lineup protocol [@GIIV],[@BCHLLSW09] from visual inference procedures is employed. The approach mirrors classical statistical inference. The procedures for doing a power comparison of competing plot designed, outlined in @GTPCCD, are followed. 
XXX reasoning for using this approach, over alternatives
eg. The lineup protocol allows a measurement of how easily participants can make the detection of a spatial pattern when two different map types are used. Data is simulated and then visualised in a way that one location shows the same significant result in the two different visualisations (hexagon and choropleth). The detection rate indicates how easily participants can make the detection with respect to the map type. Where the detection rate is the number of correct detections of the significant data pattern by participants.

In classical statistical inference hypothesis testing is conducted by comparing the value of a test statistic on a standard reference distribution, computed assuming the null hypothesis is true. If the value is extreme,  the null hypothesis is rejected, because the test statistic value is unlikely to have been so extreme if it was true. In the lineup protocol, the plot plays the role of the test statistic, and the data plot is embedded in a field of null plots. Defining the plot using a grammar of graphics [@ggplot2] makes it a functional mapping of the variables and thus, it can be considered to be a statistic. With the same data, two different plots can be considered to be competing statistics, one possibly a more powerful statistic than the other. 

To do hypothesis testing with the lineup protocol requires human evaluation. The human judge is required to identify the most different plot among the field of plots. If this corresponds to the data plot -- the test statistic -- the null hypothesis is rejected. It means that the data plot is extreme relative to the reference distribution of null plots. 

The null hypothesis is explicitly provided by the grammatical plot description. For example, if a histogram is the plot type being used, the null might be that the underlying distribution of the data is a Gaussian. Null data would be generated by simulating from a normal model, with the same mean and standard deviation as the data. In practice, the null hypothesis used is generic, such as *there is NO structure or a pattern in the plot*, and contrasted to an alternative that there is structure. 

The chance that an observer picks the data plot out of a lineup of size $m$ plots accidentally, if the null hypothesis is true is $1/m$. With $K$ observers, the probability of $k$ randomly choosing the data plot, roughly follows a binomial distribution with $p=1/m$. Fig. \ref{fig:lineup} shows a lineup of the hexagon tile map, of size $m=12$. Plot 3 is the data plot, and the remaining 11 are plots of null data. XXX Add footnote, the same data is presented in aus_cities_3_geo

```{r lineup, fig.cap = "This lineup of twelve hexagon tile map displays contains one map with a real population related structure, it is located in plot 3. The rest are null plots that contain spatial correlation between neighbours.", fig.height = 8, fig.width = 8}
ggdraw() +
  draw_plot(rasterGrob(readPNG("figures/aus_cities_3_hex.png")))
# XXX include, in the supplementary materials, this same type of figure for a choropleth lineup aus_cities_3_geo
```

In order to determine the effectiveness of a type of display, this probability is less relevant than the overall proportion of observers who pick the data plot, $k/K$. The power of the test statistic (data plot) is provided by this proportion. Power in a statistical sense is the ability of the statistic to *produce a rejection* of the null hypothesis, if it is indeed *not true*. With the same data plotted using two different displays, the display with the highest proportion of people who choose the data plot would be considered to be the most powerful statistic. 

Methodology
============

This study aims to answer two key questions around the presentation of spatial distributions:

1. Are spatial disease trends that impact highly populated small areas detected with higher accuracy, when viewed in a hexagon tile map?
2. Are people faster in detecting spatial disease trends that impact highly populated small areas when using a hexagon tile map?

Additional considerations when completing this experimental task included the difficulty experienced by participants and the certainty they had in their decision.

Australia is used for the study, with Statistical Area 3 (SA3) [@abs2016] as the geographic units. The results should apply broadly to any other geographic area of interest. XXX Specify the limits on broad application here
XXX what of areas with similar geographic v hex map displays
XXX hex map should perform similarly and not impede identification of population distributions when geo == hex
## Experimental factors

<!-- The variables changed between groups were the type of plot shown and the trend model.-->

The primary factor in the experiment is the plot type. The secondary factor is a trend model. Three trend models were developed, one mirroring a large spatial trend for which the choropleth would be expected to perform well, and two with differing level of inner city hot spots. These latter two reflect the structure seen in the thyroid cancer data (Figures \ref{fig:thyroid-choro}, \ref{fig:thyroid-hex}). This produces six treatment levels:

  - Map type: *Choropleth, Hexagon tile*
  - Trend: *South-East to North-West; Locations in three population centres; Locations in multiple population centres, *

Data is generated for each of the trend models, with four replicates, and each displayed both as a choropleth and as a hexagon tile map, which yields 12 data sets, and 24 data plots. This set of displays is divided in half, providing two sets of 12 displays, Group A and Group B. Participants were randomly allocated to Group A or B. Participants saw a data set only once, either as a choropleth or as a hexagon tile map. Table \ref{fig:exp-design} summarises the design and the allocation of the displays.


```{r exp-design, results = "asis", fig.width = 8, fig.height = 8, fig.cap = "The experimental design used in the visual inference study. XXX describe the location"}
ggdraw(xlim = c(0,1), ylim = c(0,1)) + draw_plot(rasterGrob(png::readPNG("figures/experiment_design.png")))
#XXX change names to “Hexagon Tile” and “Choropleth”
```

## Generating null data

Null data needs to be data with no (interesting) structure. In most scenarios, permutation is the main approach for generating null plots. It is used to break association between variables, while maintaining marginal distributions. This is too simple for spatial data. In spatial data, a key feature is the spatial dependence or smoothness over the landscape. To do something simple, like permute the values relative to the geographic location would produce null plots which are too chaotic, and the data plot will be recognisable for its smoothness rather than any structure of interest. 

XXX watch the language here and make the description more clear 
For spatial data, null data is stationary data, where the mean, variance and spatial dependence are constant over the geographic units. Stationary data is specified by a variogram model [@POG]. Simulating from a variogram model, where the spatial dependence is specified, generates the stationary spatial data used for the null plots.  The parameters for the Gaussian model were sill=1, range=0.3 with the variance generated by a standard normal distribution. 

The R package `gstat` [@gstat] was used to simulate 144 null sets, 12 data sets for each plot in a lineup, and 12 sets for 12 lineups.

```{r eval=FALSE, echo=FALSE}
var.g.dummy <- gstat(formula = z ~ 1, 
                     locations = ~ longitude + latitude, 
                     dummy = T, beta = 1, model = vgm(psill = 1, model = "Gau", range = 0.3),
                     nmax = 12)
```


<!-- Applying smoothing -->

The null model imposed by our hypothesis suggests that neighbours are related. The randomness induced when generating the null data was smoothed to mirror the practices employed by the Australian Cancer Atlas statisticians. 
In these 12 sets of data, each of the 12 maps were smoothed several times to replicate the spatial autocorrelation seen in cancer data sets presented in the Australian Cancer Atlas, without implementing uncertainty via transparency. XXX reference paper describing this method or describe the smoothing method here

A list of neighbours for each geographic unit was generated to use when smoothing the distributions. For each geographic unit the same spatial smoother was applied in each layer of smoothing. It kept half of the units' previous value, and derived the new half as the mean of the values of its neighbours at the previous layer of smoothing. 
XXX expressed mathematically as: 
0.5 * value of area + 0.5 MEAN(value| area is neighbour)

This smoothing allowed neighbours to be related to each other, but also allowed outliers, and showed distributions similar to the thyroid cancer distribution (Figures \ref{fig:thyroid-choro}, \ref{fig:thyroid-hex}). 


## Generating lineups


<!-- Discuss simulating the trend data -->
For each trend model, four real data displays were created by manipulating the centroid values of each of the SA3 geographic units.

The North West to South East (NW-SE) distribution was created using a linear equation of the centroid longitude and latitude values. 

The All Cities trend model was created using the distance from the centroid of each geographic unit to the closest capital city in Australia, calculated when creating the hexagon tile map using the sugarbag [@sugarbag] package.
201 of the 336 SA3s were considered greater capital city areas, the values of these areas were increased to create red clusters. The amount was chosen to make clusters around the cities visible in the choropleth display even if they were not overtly noticeable. XXX check if this is aligned to the ABS GCC areas or if not, justify

A similar selection process was applied to the Three Cities' trend model. However, for each of the four replicates for the Three Cities trend, a random sample of capital cities was taken from Sydney, Brisbane, Melbourne, Adelaide, Perth, and Hobart. Only values of the areas nearest to the three cities were increased to create clusters.


<!-- Location of data plot -->

One of the lineup locations was chosen to embed the real trend model map, in each of the four replicates, for the three trend models.
The location for the real data plot was randomly chosen from the 12 possible locations. The chance of repetition was introduced to prevent participants from deducing the remaining locations of real data plots, by process of elimination. Due to allowance re-sampling, the locations 1, 7, 10 and 11 happened to not be selected.

XXX
As seen in Fig \ref{fig:exp-design}, the choropleth and hexagon display used the same location for the real data display. To simulate a naturally occurring spatial trend, some variability was simulated for the data in all plots, using spatially correlated values. The trend model in the real plot was combined with the simulated spatially correlated values to determine the final value used to colour each area. 
Each set of lineup data was used to produce a pair choropleth map lineup and hexagon tile map lineup. These matched pairs were split between Group A and Group B according to the 2 x 3 factor experimental design depicted in \ref{fig:exp-design}.


<!-- Scaling data within a lineup -->

For each of the 144 individual maps, the values for each geographic area were rescaled to create a similar colour scale from deep blue to dark red within each map.
This meant at least one geographic unit was coloured dark blue, and at least one was red, in every map display of every lineup.

For the geographic NW-SE distribution, this resulted in the smallest values of the trend model (blue) occurring in Western Australia, the North West of Australia, and the largest values of the trend model (red) occurring in the South East. This resulted in Tasmania being coloured completely red.

For the population related displays, the clusters in the cities appeared more red than the rest of Australia.


## Analysis

### Data Cleaning


The data cleaning process considered the legitimacy of submissions and limited the results used to participants who met the standard for genuine engagement. This process involved removing data from the participants who did not provide at least three unique choices across the responses for each of the twelve lineups, and did not correctly identify the pattern in any of the 12 displays. Data was not removed if participants had made a variety of plot choices for the 12 displays, even if none were correctly detected.

XXX consider the timing of the last group above, if they got none right did they do it very quickly or actually engage?


### Descriptive statistics

Basic descriptive statistics were used to contrast the detection rate for the two types of displays. Comparison was also made across the trend models, contrasting the mean and standard detection rate for each group, who had seen the different map display type for each replicate.

Side-by-side dot plots were made of accuracy (efficiency) against map type, faceted by trend model type.

Similar plots were made of the feedback and demographic variables - reason for choice, reported difficulty, gender, age, education, having lived in Australia - against the design variables.

Plots were made in R [@RCore], with the `ggplot2` package [@ggplot2].   
   
   
### Modelling

The likelihood of detecting the data plot in the lineup can be modelled using a linear mixed effects model. 
The R [@RCore] `glmer()` function in the `lme4` [@lme4] package implements generalised linear mixed effect models. The model used includes the two main effects map type and trend model, which gives the fixed effects model to be:


$$\widehat{y_{ij}} = \mu + \tau_i + \delta_j + (\tau\delta)_{ij} + \epsilon_{i,j}, ~~~ i=1,2; ~j=1,2,3$$

where $y_{ij} = 0, 1$ is the log odds for whether the subject detected the data plot (1) or did not (0), $\mu$ is the overall mean, $\tau_i, i=1,2$ is the map type effect, $\delta_j$ is the trend model effect. We are allowing for an interaction between map type and trend model as the response is binary, so a logistic model was used.

The model specifies a logic link, this means the predicted values from the `glmer` model should be back-transformed to fit between 0 and 1. The predictions  $\widehat{p}(\eta)$ are transformed to be probabilities between 0 and 1 with the link specified below:


$$\widehat{p}(\eta) = \frac{e^{\eta}}{1 + e^{\eta}}$$ \label{eq:transform}
$$\eta = f(\tau_i,\delta_j)$$
XXX Add Random effects
 As each participant provides results from 12 lineups, this model can account for each individual participants’ abilities as it includes a subject-specific random intercept. 

XXX move these sections to before the analysis?
## Web application to collect responses

The taipan [@taipan] package for R was used to create the survey web application. 
This structure was altered to collect responses regarding participants demographics and their survey responses.
The survey app contained three tabs. Participants were first asked for their demographics, and their consent to the responses being used for analysis. The demographics collected included participants' preferred pronoun, the highest level of education achieved, their age range and whether they had lived in Australia. 

For each of the twelve displays every participant was asked three questions; their plot choice, the reason for their choice and the difficulty or certainty of deciding for each plot. The time taken by the participant to view each lineup and respond to these questions was recorded.

<!-- When participants completed the twelve evaluations the survey application triggered a data analysis script. This created a data set with one row per evaluation. Containing the responses to the three questions. The script also added the title of the image, which indicated the type of map display, the type of distribution hidden in the lineup, and the location of the data plot. It also calculated the time taken by participant to view each lineup. -->

Each participant used the internet to access the survey.
The data transfer from the web application to the data set took place using a secure link to the googlesheet used to store results. The application connected to the googlesheet using the googlesheets [@sheets] R package when participants opened the application, and interacted again when participants chose to submit the survey. At this time it added the participant's responses to the twelve lineup displays as twelve rows of data in the googlesheet.

## Participants

Participants were recruited from the Figure Eight crowd sourcing platform [@figeight] to evaluate lineups.
The lineup protocol expects that the participants are uninvolved judges with no prior knowledge of the data, to avoid inadvertently affecting results. Potential participants needed to have achieved level 2 or level 3 from prior work on the platform. All participants were at least 18 years old.

Participants were allocated to either group A or group B when they proceeded to the survey web application. There were 92 participants involved in the study. All participants read introductory materials, and were trained using three test displays, to orient them to the evaluation task. All participants who completed the task were compensated $AUD5 for their time, via the Figure Eight payment system.

A pilot study was conducted in the working group of the Econometrics and Business Statistics Department of Monash University. This allowed us to estimate the effect size, and thus decide on number of participants to collect responses from.

## Demographic data collection

Each participant answered demographic questions and provided consent before evaluating the lineups.

Demographics were collected regarding the study participants:

 - Gender (female / male / other),
 - Education level achieved (high school / bachelors / masters / doctorate / other),
 - Age range (18-24 / 25-34 / 35-44 / 45-54 / 55+ / other)
 - Lived at least for one year in Australia (Yes / No )

Participants then moved to the evaluation phase.
The set of images differed for Group A and Group B.
After being allocated to a group, each individual was shown the 12 displays in randomised order.



<!--
## Experimental design


The choropleth map was used as the comparative visualisation for presenting the lineups [@VVSIALM] as this is the common display for spatial cancer data.
Geographic distributions usually have some degree of spatial autocorrelation between neighbours.
This feature was incorporated in all maps shown in the lineup displays, the map that contained the trend feature shown in only one set of data was also affected by spatial autocorrelation.
A line up protocol was implemented to arrange $N$ maps in each display. 
A reasonable amount of null plots $N-1$ in the lineup was chosen to ensure the real data map was well hidden. A reasonable number of plots to show in each lineup, $N = 12$ was chosen to not overwhelm participants due to the detailed choropleth maps of Australian SA3 areas.

The hypotheses for each lineup are
$H_0$ : All plots look the same
$H_a$ : One plot looks different to the other plots 


The same data was visualized on a choropleth map, and on a hexagon tile map, however in this study the participants only saw one of these two displays. The accuracy and times taken were aggregated for Groups and.
Comparing the results of participants who see the choropleth to those who see a hexagon tile map will show that population related distributions are spotted more frequently in a hexagon tile map display.


Let $n$ be the number of independent observers and $x_i$ the
number of observers who picked plot $i$, $i = \{1,...,m\}$

Then $x_i, x_2, ..., x_m$ follows a multinomial
distribution$Mult_{\pi_1, \pi_2, ...., \pi_m}(x_i, x_2, ..., x_m)$ with 
$\sum_i \pi_i = 1$, where $\pi_i$ is the probability that plot $i$ is picked by an
observer, which we can estimate as $\hat{\pi}_i = x_i/n$.
The researchers compared the length of time taken, and the accuracy of the participants choices.
The power of a lineup can therefore be estimated as the ratio of correct
identifications $x$ out of $n$ viewings.

## The methods of data analysis used
The data analysis methods used in order to analyse and collate the results included downloading the survey submissions and opening them into the analysis software R [@RCore].
   
For each of the 12 lineup displays the researchers calculated:

 - accuracy: the proportion of subjects who detected the data plot
 - efficiency: average time taken to respond
-->


Results
============

<!--
The results section explores the [Accuracy] and contrasts the detection rate of the real data plot in the lineups.
The second section considers the [Speed] participants had when submitting their responses to the three questions asked regarding each display.
The [Certainty] participants felt about their decisions and the [Reason] for their choice in plot is also discussed.
-->


Responses were collected from 92 participants, who each evaluated 12 lineups. Set A was evaluated by 42 participants, and 53 evaluated set B. 
Data was collected but not included in the analysis for five participants. Responses were analysed for accuracy and speed. The certainty in decision and reasons of subjects answers is also examined. 
XXXAre there differences between group A and B? This would be a good place to show that randomization is working. Underlying assumption: there a no group specific influences on test results.

## Participant demographics

Of the 92 participants, 67 were male, and 25 female. Most participants (56) had a Bachelors degree, 13 had a Masters degree, and the remaining 23 had high school diplomas.

## Accuracy

Fig. \ref{fig:detect-compare} displays the average detection rates for the two types of plot separately for each trend model. Each trend model was tested using four repetitions, evaluations on the same data set were seen as either choropleths or hexagon tile maps by each group as specified in Table. \ref{fig:exp-design}; the detection rates for each display are connected by a line segment. The Three Cities and All Cities trend models shown in the hexagon tile map allowed viewers to detect the data plot substantially more often than the choropleth counterparts. 

One replicate for the All Cities group had similar detection rates for both plot types, the rate of detection using the choropleth map was much higher than other replicates.XXX potentially explain why
Surprisingly, participants could also detect the gradual spatial trend in the NW-SE group from the hexagon tile map. We expected that the choropleth map would be superior for the type of spatial pattern, but the data suggests the hexagon tile map performs slightly better, or equally as well.

XXX overall detection rates
While there is clearly a difference between map types (Page 12, Lines 306-316), the overall level of detection rates is low. I would expect some discussion of these low results. Are these a result of the data simulation process (i.e., the “true” signal might not be different enough)? How do the results compare to the expected detection rates if selected by chance, as discussed in Section 2.2?

```{r detect-compare, fig.cap = "The detection rates achieved by participants are contrasted when viewing the four replicates of the three trend models. Each point shows the probability of detection for the lineup display, the facets separate the trend models hidden in the lineup. The points for the same data set shown in a choroleth or hexagon tile map display are linked to show the difference in the detection rate.", fig.height=5}
# Detection rate for each lineup (image)
d_smry <- d %>% group_by(trend, type, replicate) %>%
  # pdetect measures the aggregated accuracy of the choices
  summarise(pdetect = length(detect[detect == 1])/length(detect)) %>%
  ungroup()

# Numerical summary
diffs <- d_smry %>% spread(type, pdetect) %>%
  mutate(dif = `Hex.` - `Choro.`)

# Plot summary
ggplot(d_smry, aes(x = type, y = pdetect, color = trend)) +
  geom_point(size = 2) +  
  geom_line(size = 1, aes(group = replicate)) +  
  facet_wrap(~trend) +
  scale_color_manual(values = trend_colors) +
  xlab("Type of areas visualized") +
  ylab("Detection rate") + 
  ylim(0,1) +
  guides(scale = "none")
# XXX remove legend
# use color to distinguish between replicates (facets and color for trend models is redundant)
```


Table. \ref{tab:desc-stats} shows the means and standard deviations of the detection rate for each type of plot and each trend model. This also gives the standard deviations, the smallest standard deviation for all sets of replicates was the Three Cities trend model shown in a Choropleth display. This group of displays had a very small detection rate of 0.04. The mean detection rate for the Three Cities trend model shown as choropleth map lineups was also the smallest at 0.40.
The geographic trend was expected to be detected equally in the two displays, however North-West to South-East (NW-SE) trend model had a higher mean detection rate for the hexagon tile map displays, but the difference in the means of detection rate was only 0.10.


```{r desc-stats, results = "asis"}
types <- c("Choro.", "", "Hex.", "")

d %>% group_by(type, trend) %>%
  summarise(m = as.character(round(mean(detect), 2)),
      std.dev = as.character(round(sd(detect), 2))) %>% 
  mutate(std.dev = ifelse(std.dev == 0.5, "(0.50)", paste0("(", std.dev, ")"))) %>%   mutate(m = ifelse(m == 0.4, "0.40", m)) %>% 
  gather(stat, value, m, std.dev) %>% 
  pivot_wider(names_from = "trend", values_from = "value") %>% 
  arrange(type) %>% ungroup() %>% 
  mutate(Type = types) %>% select(Type, `NW-SE`, `Three Cities`, `All Cities`) %>%
  knitr::kable(., format = "latex", align = "lccc", booktabs = TRUE, 
    linesep = c("", "\\addlinespace"),
    caption = "The mean and standard deviation of the rate of detection for each trend model, calculated for the choropleth and hexagon tile map displays.")
# XXX remove legend
```


Table. \ref{tab:detect-glmer1} presents a summary of the generalised linear mixed effects model, testing the effect of plot type and trend model on the detection rate. The results support the summary from Fig. \ref{fig:detect-compare} and all parameters are statistically significant despite the large standard deviations observed in Table. \ref{tab:desc-stats}. Overall, the hexagon tile map performs marginally better than the choropleth for all trend models. 

XXX rephrased this sentence, when is it better/worse?
Population related distributions are more easily detected using a hexagon tile map display compared to when a choropleth map lineup was used.

XXX check the numbers on the calculations to make it possible to replicate the detection probabilities mentioned in the text,
The log odds of detection show in Table. \ref{tab:detect-glmer1} can be back transformed after taking the sum of all terms for the trend and type of display that are of interest.
For the NW-SE distribution, the predicted detection rate for the hexagon tile map display increases the predicted probability of detection to `r round(exp(0.46+0.07)/(1+exp(0.46+0.07)), 2)` from `r round(exp(0.07)/(1+exp(0.07)),2)` for choropleths, this is almost exactly the difference seen in the table of means and is significant only at the 0.05 level. 

When a choropleth map display is used, the predicted detection rate for the Three Cities trend, `r round(exp(0.07-3.41)/(1+exp(0.07-3.41)),2)`; this is extremely low, especially compared to the NW-SE trend of `r round(exp(0.07)/(1+exp(0.07)),2)`.
When the All Cities trend is presented in a choropleth display the predicted probability of detection is `r round(exp(0.07-1.34)/(1+exp(0.07-1.34)),2)`.
The hexagon tile map has a substantially high detection rate for the display of a Three Cities trend `r round(exp(0.46+0.07-3.41+2.44)/(1+exp(0.46+0.07-3.41+2.44)),2)` and All Cities trend `r round(exp(0.46+0.07-1.34+1.16)/(1+exp(0.46+0.07-1.34+1.16)),2)`.


```{r detect-glmer1, results="asis"}
# Mixed effects models
glmer1 <- glmer(detect ~ type*trend + (1|pid), 
              family = binomial, data = d)

glmer_terms <- c("Intercept", "Hex.", "Three Cities", "All Cities",
  "Hex:Three Cities", "Hex:All Cities")

detection_rates <- tidy(glmer1) %>%
  mutate(detection_rates = round(exp(estimate)/(1+exp(estimate)),2)) %>%
  select(term, estimate, detection_rates) %>% pull(detection_rates)

tidy(glmer1) %>%
  mutate_at(.vars = c("estimate", "std.error"), round, 2) %>% 
  mutate(p.value = round(p.value, digits=2)) %>% 
  rowwise() %>%
  mutate(sig = case_when(
  p.value <= 0.001 ~ "$^{***}$",
  p.value <= 0.01 ~  "$^{**}$",
  p.value <= 0.05 ~  "$^{*}$",
  p.value <= 0.01 ~  "$^{.}$",
  TRUE ~ "$^{ }$")) %>%
  ungroup() %>% 
  filter(!is.na(std.error)) %>%
  mutate(term = glmer_terms) %>% 
  select(Term = term, 
    Est. = estimate, 
    Sig. = sig,
    `Std. Error` = std.error, 
    `P val` = p.value) %>% 
  knitr::kable(format = "latex", escape = FALSE, align= "rrlrr", 
    booktabs = T, linesep = c("", "\\addlinespace", "", "\\addlinespace", ""), 
    caption = "The model output for the generalised linear mixed effect model for detection rate. This model considers the type of display, the trend model hidden in the data plot, and accounts for participant performance.")
# XXX remove legend
```

XXX see results under control of socio demographics. A second model including those variables could give some hints (direct effects, reduction intra-participant variance).  For reference see: (Vanderplas & Hofmann, 2016)

## Speed

Fig. \ref{fig:beeswarm} shows horizontally jittered dot plots to contrast the time taken by participants to evaluate each lineup when viewing each type of display. The time are also separated by trend model and whether the data plot was detected or not detected. The time taken to complete an evaluation ranged from milliseconds to 60 seconds. The average time taken for type of display is shown as a large coloured dot on each plot. when considering the heights of the green and orange dots, there is little difference in the average time taken to read a choropleth or hexagon tile map. Comparing the same coloured dot across each trend model row, there is a slight increase in the time taken to correctly detected the data plot in the hexagon tile map lineup, but little difference in evaluation time for the choropleth display. However, there were substantially less correct detections for choropleth lineups for the Three cities and All Cities trends.

XXX There is a clear gap between these participants and those who took longer to complete the experiment. Does the accuracy differ between these two groups? If the accuracy is significantly lower for the quicker participants, this might indicate that they chose a location by chance to finish the task as quickly as possible. What is the experience from the pilot study? Are these times taken plausible? If not, one might consider removing those cases with implausibly short decision times from the sample to eliminate randomly chosen locations. Regardless of how you deal with these cases, this detail should be discussed in the text.

```{r beeswarm, fig.cap = "The distribution of the time taken (seconds) to submit a response for each combination of trend, whether the data plot was detected, and type of display, shown using horizontally jittered dotplots. The colored point indicates average time taken for each plot type. Although some participants take just a few seconds per evaluation, and some take as much as 60 seconds, but there is very little difference in time taken between plot types.", fig.height=8}
# Di playing
s <- d %>% group_by(type, trend, detect_f) %>%
  summarise(m=median(time_taken), 
            q1=quantile(time_taken, 0.25), 
            q3=quantile(time_taken, 0.75))
library(ggbeeswarm)
ggplot() + 
  geom_quasirandom(data=d, aes(x=type, y=time_taken), alpha=0.9) + 
  #geom_hline(data=s, aes(yintercept=m, colour=type)) +
  geom_point(data=s, aes(x=type, y=m, color=type), size=5, alpha=0.7) +
  #geom_errorbar(data=s, aes(x=type, ymin=q1, ymax=q3, colour=type), width=0.3, size=2) +
  scale_color_brewer("", palette = "Dark2") +
  facet_grid(trend~detect_f) + 
  ylab("Time taken (seconds)") + xlab("") +
  theme(legend.position="bottom")
```


## Certainty

Participants provided their level of certainty regarding their choice using a five point scale. 
Unlike the accuracy and speed of responses that were derived during the data processing phase, this was a subjective
assessment by the participant prompted by the question: ‘How certain are you about your choice?’.
Fig. \ref{fig:certainty} shows the amount of times participants provided each level of certainty. This was separated for each combination of trend models and display type, and coloured depending on whether a participant correctly detected the data plot in the lineup.
Participants often chose 4 or 5 when viewing the population related trends in the choropelth display, even though they were often incorrect when viewing an All Cities trend and overwhelmingly incorrect for the Three Cities trend. This shows overconfidence in their detection ability when using a choropleth map display. Participants were less likely to be certain when their choice was incorrect and they were viewing a hexagon tile map.
For each trend model, participants were more likely to doubt their choice and choose 1 or 2 in the hexagon tile map displays, even though many had made the correct choice. 


```{r certainty, fig.cap = "The amount of times each level of certainty was chosen by participants when viewing hexagon tile map or choropleth displays. Participants were more likely to choose a high certainty when considering a Choropleth map. The mid value of 3 was the default certainty, it was chosen most for the Hexagon tile map displays.", fig.height=5}
d <- d %>% 
  mutate(certainty = as_factor(certainty)) %>% 
  mutate(replicate_f = as_factor(replicate)) 
 
d %>% 
  mutate(Detected = factor(detect_f, 
    levels = c("Detected? Yes", "Detected? No"), 
    labels = c("Yes", "No"))) %>% 
ggplot(aes(x = certainty, fill = Detected)) +  
  scale_fill_manual(values = detect_f_colors) +
  geom_bar() + facet_grid(type ~ trend) +
  theme(legend.position = "bottom") + xlab("Level of certainty")
```


## Reason

XXX merge with accuracy section above and link to table 3 
Participants were asked why they had made their plot choice and were able to select from a set of suggested reasons. 
"Colour trend across the areas" was the most common selection for NW-SE trend displays.

The reasons chosen by participants from the list provided to them varied more when viewing choropleth displays than the hexagon tile map.
The hexagon tile map displays resulted in "Clusters of colour" as the most common choice made by participants.

The choice "None of these reasons" was used as the default value to minimise noise from participants who did not select a response. 

```{r reason, results = "asis"}
# Qualitative analysis of reason
d %>% 
  mutate(reason = ifelse(reason =="0.0", "no reason", reason)) %>% 
  mutate(Detected = ifelse(detect_f == "Detected? Yes", "Yes", "No"),
    Trend = trend) %>% 
  group_by(Trend, Detected, type) %>% 
  count(reason) %>% 
  mutate(prop = round(n/sum(n), 2), r_prop = paste0(reason, ":", prop)) %>% 
  top_n(1, n) %>% summarise(reasons = paste(reason, collapse=", ")) %>% 
  pivot_wider(names_from = c("type"), values_from = c("reasons")) %>% 
  ungroup() %>% 
  knitr::kable(., format = "latex", booktabs = TRUE,
    linesep = c("", "\\addlinespace", "", "\\addlinespace",""),
    caption = "The amount of participants that selected each reason for their choice of plot when looking at each trend model shown in Choropleth and Hexagon Tile maps. The facets show whether or not the choice was correct.") %>% 
  #kable_styling(bootstrap_options = "hold_position") %>% 
  collapse_rows(., columns = 1)
```



Discussion
============

The intention of this study was to contrast the use of the choropleth map and the hexagon tile map.
The visual inference lineup protocol was employed to contrast the effectiveness of the displays.
The results have shown that overall the use of the hexagon tile map display allows participants to find the data plot in the lineup more often.
Using the visual inference protocol this result can be extended to show that it is a valid alternative display to communicate spatial distributions of population related data.

We expected that the choropleth map would be superior for communicating the spatial pattern of geographic distributions. The data suggest that the participants perform slightly better or equally as well for each replicate in each trend model across the two displays. Table II shows that the difference in the mean detection rate for the two trend models was 0.10.

The differences seen in the \ref{fig:detect-compare} plot and Table. \ref{tab:desc-stats} are reflected in the model results. Surprisingly the difference for the geographic distribution was significant at the 0.05 level.
It also showed that the hexagon tile map display performs marginally better than the choropleth for all trend models. Unexpectedly the detection rate suffers when using a choropleth map to display population related distributions.

While the significance of the difference in detection was the key focus of this experiment, the secondary focus was the time taken by participants. it was expected that the participants may take longer to consider the hexagon tile map distribution but would be able to detect the data plot in the lineup.
The bimodal distributions seen in Fig. \ref{fig:beeswarm} showed very little difference in the mean evaluation times. As the maximum time of all of the distributions approached 60 seconds it cannot be said that the participants' took longer to evaluate the hexagon tile map displays. 

The responses to the questions asked of participants included the reason for their choice and the certainty around their choice.
Fig. \ref{tab:reason} shows high levels of certainty of 4 and 5 were chosen by participants when looking at the population distributions in a choropleth map display show that they were over confident when attempting to find the real data plot in the choropleth map displays. Participants performed better on the NW-SE distribution shown in the choropleth display and were reasonably confident about their decisions.
The high levels of the mid range value of 3 could indicate that the participant did not want to provide a response, as this was the default value. Those who chose level 4 or 5 were equally likely to be correct for the three cities lineups, but more likely to be correct than incorrect for the other two trend models.


The colour scaling applied in Three cities and All cities displays resulted in the rural areas of the real data plot appearing more blue or yellow than the other plots in the lineups.
Due to the consistent colouring of rural areas in a choropleth display, the choice "All areas have similar colours" was most common reason for a participants choice. The All Cities displays coloured the inner-city areas of all capital cities more red, this was observable to participants and explains the equal choice of the city clusters or rural colour consistency. 
Choosing "Clusters of colour" was expected when participants viewed the Hexagon tile map display of the All Cities and Three Cities distributions. It was unexpected that it was also the most common reason for the NW-SE hexagon tile map displays. 
Due to the spatial covariance introduced in the smoothing, groups of similarly coloured hexagons were present in all of the hexagon tile map displays. All Cities and Three Cities distributions of real data trends had distinctly different patterns or red inner-city areas, while some of the plots in each lineup may have shared similar features.

XXX Add limitations - of the method used, and of application to other sets of  geographic areas
XXX extrapolate the results to other modelled outputs beyond cancer statistics to obtain broader coverage. There is considerable value in using a hexagon tile map for showing statistics and modelled output in statistical area levels across Australia or similar.

XXX Add discussion of future work, related to this study.

Conclusion
============

The choropleth map display and the tessellated hexagon tile map have been contrasted using the lineup protocol. The hexagon tile map was significantly more effective for spotting a real population related data trend model hidden in a lineup.

The hexagon tile map display should be considered as an alternative visualisation method when communicating distributions that relate to the population across a set of geographic units. As an additional display to the familiar choropleth map, cancer atlas products may benefit from the opportunity to allow exploration via an alternative display. The spatial distributions used to test these displays were inspired by the real spatially smoothed estimates of the cancer burden on Australian communities. However, this technique may be extended to other population related distributions, such as other diseases.

The increasing population densities of capital cities despite large land area exacerbates the difference in the smallest and largest communities.
The population density structure of Australia can be considered similar to that of Canada, New Zealand and many other countries. Therefore, this display is not only relevant to Australia, but all nations or population distributions that experience densely populated cities separated by vast rural expanses.


Acknowledgment
==============

The authors would like to thank the Australian Cancer Atlas team for discussions regarding alternative spatial visualisations, and Professor Kerrie Mengersen and Dr Earl Duncan for regular meetings filled with suggestions and comments. Mitchell O'Hara-Wild was a co-developer of the taipan [@taipan] R package for image tagging, used as the base for the web app constructed to collect participant evaluations of lineups. We are thankful for the NUMBATs (Non-Uniform Monash Business Analytics Team) for participating in the pilot study that helped to assess the experimental design and determine an appropriate sample size for the study. 

XXX Decide where this information belongs
==============

The source code to produce this document can be found on [GitHub](https://github.com/srkobakian/experiment/paper).
Supplementary materials have been included to discuss the survey procedures and the lineups that were used.
The full set of images can be found here, too. 

The supplementary material contains:

- Additional analysis of the experimental results
- Survey procedure including training materials for the participants
- 24 lineups as images, that were used in the experiment
- 12 data sets used to construct the lineups

The analysis of the work was completed in R [@RCore] with the use of the following packages:

 - For document creation: rmarkdown [@rmarkdown], rticles [@rticles], knitr [@knitr].
 - For lineup creation and data analysis: tidyverse [@tidyverse], nullabor [@nullabor], ggthemes [@ggthemes], RColorBrewer [@RColorBrewer].
 - For image displays: cowplot [@cowplot], png [@png], grid [@grid].
 - For modelling and presentation of models: gstat [@gstat], lme4 [@lme4], kableExtra [@kableExtra].

Ethics approval for the online survey was granted by QUT’s Ethics Committee (Ethics Application Number: 1900000991). All applicants provided informed consent in line with QUT regulations prior to participating in this research. 

References {.numbered}
==========
