---
title: "Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics"
affiliation:
  author-columnar: true         ## one column per author
  #institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields
  institution:
    - name: Queensland University of Technology
      department: Science and Engineering Faculty
      location: Brisbane, Australia
      email: stephanie.kobakian@qut.edu.au
      mark: 1
      author:
        - name: Stephanie Kobakian
    - name: Monash University
      department: Econometrics and Business Statistics Faculty
      location: Melbourne, Australia
      email: dicook@monash.edu
      mark: 1
      author:
        - name: Dianne Cook
keywords: ["statistics", "visual inference", "geospatial", "population"]
abstract: |
  The choropleth map display is commonly used for communicating spatial distributions across geographic areas. However, when choropleths are used, the sizes of areas can lead to the misinterpretation of the distribution. The visualization method used to present geospatial data will influence the understanding of the distribution derived by map users. Choosing an effective alternative display could positively influence the communication of the spatial distribution. The hexagon tile map is presented as an alternative display for visualizing population related distributions effectively. Visual inference is used to measure the power of design, and the choropleth is used as a comparison. The hexagon tile map display is also tested using a distribution that is directly related to the geography, with values monotonically decreasing from the North-West to South-East areas of Australia. This study finds the single map in a hexagon tile map lineup that contains a population related distribution is detected with greater probability than the same data displayed in a choropleth map. These findings should encourage map creators to implement alternative displays and consider a hexagon tile map when presenting spatial distributions of heterogeneous areas.
dev: pdf
bibliography: paper.bib
output:
  bookdown::pdf_book: 
      base_format: rticles::ieee_article
      number_sections: true
editor_options: 
  chunk_output_type: inline
link-citations: true
citation_package: biblatex
csl: springer-vancouver.csl
header-includes: \usepackage{graphicx}
---

<!-- Regular paper – 12 double column pages (Submissions may be up to 18 pages in length, subject to MOPC. All regular paper page limits include references and author biographies.) -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  error = FALSE, 
  message = FALSE,
  cache = FALSE,
  dpi = 300)
```

```{r libraries}
# Load Libraries
library(tidyverse)
library(readxl)
library(broom)
library(cowplot)
library(png)
library(grid)
library(lme4)
library(ggthemes)
library(RColorBrewer)
library(knitr)
library(kableExtra)

options(knitr.kable.digits = "2")
```

```{r data}
trend_colors <- c(
  "NW-SE" = "#B2DF8A",
  "Three Cities" = "#A6CEE3",
  "All Cities" = "#1F78B4")
  
type_colors <- c(
  "Choro." = "#fcae91",
  "Hex." = "#a50f15")

detect_f_colors <- c(
  "No" = "#66C2A5",
  "Yes" = "#FC8D62")

detect_colors <- c(
  "Detected? No" = "#66C2A5",
  "Detected? Yes" = "#FC8D62")

  # Downloaded data
d <- read_xlsx("data/experiment-export.xlsx", sheet=2) %>%
  filter(!is.na(contributor)) %>%
  mutate(contributor = factor(contributor))

# Check data set 
# Need to clean multiple entries, 48, 24
# remove duplicated entries due to submit button
d <- d %>% group_by(group, contributor, image_name) %>%
  slice(1) %>% ungroup() %>% 
  arrange(group, contributor, plot_order)

# Remove contributors who did not provide answers to most questions
keep <- d %>% count(contributor, sort = TRUE) %>% filter(n > 10)
d <- d %>% 
  filter(contributor %in% keep$contributor) %>%
  filter(contributor != "1234567890")

# Remove contributors who did not provide any choices
bad_contribs <- d %>% group_by(contributor) %>% 
  summarise(sum0 = sum(choice)) %>% 
  filter(sum0 == 0) %>% 
  pull(contributor)

d <- d %>% 
  filter(!(contributor %in% bad_contribs))


n_contributors <- d %>% count(contributor, sort=TRUE) %>% 
  summarise(n_contributors = length(contributor))

d <- d %>% mutate(certainty = factor(as.character(certainty),
  levels = c("1", "2", "3", "4","5"), ordered=TRUE))
```


```{r reps}
replicate <- tibble(image_name = c("aus_cities_12_geo.png", "aus_cities_12_hex.png", 
                                   "aus_cities_3_geo.png", "aus_cities_3_hex.png",
                                   "aus_cities_4_geo.png", "aus_cities_4_hex.png",
                                   "aus_cities_9_geo.png", "aus_cities_9_hex.png",
                                   "aus_nwse_2_geo.png", "aus_nwse_2_hex.png",
                                   "aus_nwse_3_geo.png", "aus_nwse_3_hex.png",
                                   "aus_nwse_5_geo.png", "aus_nwse_5_hex.png",
                                   "aus_nwse_6_geo.png", "aus_nwse_6_hex.png",
                                   "aus_three_12_geo.png", "aus_three_12_hex.png",
                                   "aus_three_5_geo.png", "aus_three_5_hex.png",
                                   "aus_three_8_geo.png", "aus_three_8_hex.png",
                                   "aus_three_9_geo.png", "aus_three_9_hex.png"),
                    replicate = c(1, 1, 2, 2, 3, 3, 4, 4, 
                                  1, 1, 2, 2, 3, 3, 4, 4,
                                  1, 1, 2, 2, 3, 3, 4, 4))
# Add rep info to data
d <- d %>% left_join(., replicate, by = "image_name")
```

```{r pdetection_group}
# Tidy for analysis
d <- d %>% 
  separate(image_name, c("nothing", "trend", "location", "type", "extra"), remove = FALSE) %>%
  select(-nothing, -extra) %>%
  mutate(location = as.numeric(location), 
    # detect measures the accuracy of the choice
         detect = ifelse(location == choice, 1, 0)) %>% 
  mutate(trend = case_when(
    trend == "nwse" ~ "NW-SE",
    trend == "cities" ~ "All Cities",
    trend == "three" ~ "Three Cities")) %>% 
  mutate(trend = fct_relevel(trend, "NW-SE","Three Cities","All Cities")) %>% 
  mutate(type = case_when(
    type == "hex" ~"Hex.",
    TRUE~"Choro.")) %>% 
    mutate(detect_f = factor(detect, levels = c(0,1), labels = c("Detected? No", "Detected? Yes")))

plots <- d %>% group_by(group, trend, type, location) %>%
  # pdetect measures the aggregated accuracy of the choices
  summarise(pdetect = length(detect[detect == 1])/length(detect)) 
```


Introduction
=============
<!-- General: comparison of displays, motivate hexagon tile map, enough info to motivate aims -->
<!-- geographies on choropleth -->
This study compares the effectiveness of a new type of display, a hexagon tile map, against the standard, a choropleth map, for communicating information about disease statistics. The choropleth map is the traditional  approach for visualizing aggregated statistics across administrative boundaries. A hexagon tile map forgoes the familiar boundaries, in favor of representing each geographic unit as an equally sized hexagon, placed approximately in the correct spatial location. The hexagon tile map builds on existing displays, such as the cartogram, and tessellated hexagon displays. It differs in that it relaxes the requirement to have connected hexagons, and allows sparsely located hexagons. The algorithm to construct a hexagon tile map is available in the R package sugarbag [@sugarbag]. This type of display may be useful for other countries, and other purposes. 

The hexagon tile map was designed for Australia, motivated by a need to display spatial statistics for a new Australian Cancer Atlas. None of the existing approaches for creating cartograms or hexagon tiling perform well for the Australian landscape, which has vast open spaces and concentrations of population in small regions clustered on the coastlines. 

The Australian Cancer Atlas [@atlas] is an online interactive web tool created to explore the burden of cancer on Australian communities. There are many cancer types to be explored individually or aggregated. The Australian Cancer Atlas allows users to explore the patterns in the distributions of cancer statistics over the geographic space of Australia. It uses a choropleth map display and diverging color scheme to draw attention to relationships between neighboring areas. The hexagon tile map may be a useful alternative display for the atlas. 

The experiment was conducted using the lineup protocol, a visual inference procedure [@GIIV], to objectively test the effectiveness of the two displays. 

The paper is organised as follows. The next section discusses the background of geographic data display and visual inference procedures. The [Methodology] Section describes the methods for conducting the experiment and analysing the results. The results are summarized in Section [Results]. 
Background
=============

## Spatial data displays

<!-- foundation, tradition -->  
Spatial visualisations communicate the distribution of statistics over geographic landscapes. The choropleth map [@EI], [@BCM] is a traditional display. It is used to present statistics that have been aggregated on geographic units. Creating a choropleth map involves drawing polygons representing the administrative boundaries, and filling with colour mapped to the value of the statistic. The choropleth map places the statistic in the context of the spatial domain, so that the reader can see whether there are spatial trends, clusters or anomalies. This is important for digesting disease patterns. If there is a trend it may imply that the disease is spreading from one location to another. If there is a cluster, or an anomaly, there may be a localized outbreak of the disease. Aggregating the statistic on administrative units, provides a level of privacy to individuals, while allowing the impact of the disease on the community to be analyzed.

<!-- Many sets of Australian statistical areas have a large disparity between the smallest and largest units.
The rural communities in Australia operate on a much larger geographic space than small inner-city communities. When a spatial distribution is related to the size of the areas, or the population density, the size of the regions can allow erroneous conclusions to be drawn about the state of the statistic over the entire population, when presenting the statistics on a geographic map base. This occurs as large regions filled with a consistent colour or pattern can draw the attention of map readers, and small regions are not paid equal attention [@CTTMB]. An alternative display can be used to honestly communicate a spatial distribution for a set of heterogeneous areas [@NISCC]. 
This work aims to show that a hexagon tile map display is a viable alternative to the geographic map base for presenting population statistics.


A choropleth map is not the only display that can be used for presenting geospatial data. 
Alternative maps include various cartograms and tessellated tile maps. They allow other variables to be incorporated in the display to highlight the statistical values of various geographic areas.
-->

```{r liver, fig.height = 6, fig.width = 4, fig.cap = "A choropleth map (a), and hexagon tile map (b) of the smoothed average of liver cancer diagnoses for Australian males. The diverging colour scheme uses dark blue areas for much lower than average diagnoses, yellow areas with diagnoses around the Australian average, red shows diagnoses much higher than average. The hexagon tile map shows concentrations of higher than expected liver cancer rates in the cities of Melbourne and Sydney, which is not visible from the choropleth.", results = "asis"}
ggdraw(ylim = c(0,2), xlim = c(0,1.3), clip = "on") +
  draw_plot(rasterGrob(readPNG("figures/aus_liver_m.png")), -0.15, 0.9, 1.6, 1.2) +
  draw_plot(rasterGrob(readPNG("figures/aus_liver_m_hex.png")), -0.15, -0.1, 1.6, 1.2) +
  draw_plot_label(c("a", "b"), 
    c(0.05, 0.05), c(1.95, 1), size = 20, color = "white") + 
  theme(plot.background = element_rect(fill = "black"))
```
<!--
This will create the figure as needed in the paper, but I can't get the width argument to work properly.
\begin{figure*}[!t]
\centering
\includegraphics{paper_files/figure-latex/liver-1.pdf}
\caption{A choropleth map (a), and hexagon tile map (b) of the smoothed average of liver cancer diagnoses for Australian males. The diverging colour scheme uses dark blue areas for much lower than average diagnoses, yellow areas with diagnoses around the Australian average, red shows diagnoses much higher than average. The hexagon tile map shows concentrations of higher than expected liver cancer rates in the cities of Melbourne and Sydney, which is not visible from the choropleth.}
\label{fig:liver2}
\end{figure*}
-->

The choropleth map is effective if the size of the geographic units is relatively uniform. This is not the case for most countries. Size heterogeneity in administrative units is particularly extreme in Australia: most of the landscape of Australia is sparsely settled, with the population densely clustered into the narrow coastal strips. A choropleth map focuses attention on the geography, and for heterogeneously sized areas it presents a biased view of the population related distribution of the statistic [@CBATCC]. *Land doesn't get cancer, people do* -- a more effective way to communicate the spatial distributions of cancer statistics is needed. 

<!-- Cartogram -->
A cartogram is a general solution for better displaying a population-based statistic. It transforms the geographic map base to reflect the population in the geographic region, while preserving some aspects of the geographic location. There are several cartogram algorithms [@ACTUC], [@CBATCC]; each involves shifting the boundaries of geographic units, using the value of the statistic to increase or decrease the area taken by the geographic unit on the map. The changes to the boundaries result in cartograms that accurately communicate population by map area for each of the geographic units but can result in losing the familiar geographic information. For Australia, the transformations warp the country so that it is no longer recognizable.

<!-- Other cartograms -->
Alternative algorithms make various trade offs between familiar shapes and representation of geographic units. The non-contiguous cartogram method [@NAC] keeps the shapes of geographic units intact, and changes the size of the shape. This method disconnects areas creating empty space on the display losing the continuity of the spatial display of the statistic. The Dorling cartogram [@ACTUC] represents each unit as a circle, sized according to the value of the statistic. The neighbour relationships are mostly maintained by how the circles touch. A similar approach was pioneered by Raisz [-@RSCW], using rectangles that tile to align borders of neighbours [@CDWCS]. A thorough review of the array of methods, as suitable for cancer atlas displays is available in XXX.

<!-- Hexagon tile map -->
The hexagon tile map algorithm, automatically matches spatial regions to their nearest hexagon tile, from a grid of tiles. It has the effect of spreading out the inner city areas while maintaining the spatial locations or regions in remote areas. The algorithm is available in the R package, sugarbag [@sugarbag]. Figure \ref{fig:liver} shows the hexagon tile map, along with the choropleth map of liver cancer rates in Australia. Colour maps from substantially below average (blue) to substantially above average (red) rates. The inner city areas are expanded out, making it possible to see the cancer incidence in the small, densely populated areas. Remote regions are represented by isolated hexagons, which is not ideal, but maintains the spatial location of these data values. It is of interest to know how well the spatial distribution is perceived for this display, in comparison to the choropleth.

<!-- Specific problem for Australia 
Australia is an extreme case of heterogeneous geographic units, with the large differences between the smallest and largest geographic units across various granularities.
To communicate information regarding the small inner-city geographic units of Australian cities, an alternative display must be employed. This will emphasize the value of the statistic for areas with a high population. Fig. \ref{fig:liver} shows the distribution of the smoothed average amount of liver cancer diagnoses for Australian Males from 2005 to 2014.
These inner-city areas are not visible on the choropleth map (a). The hexagon tile map (b) shows orange hexagons for the inner-city SA2s with higher than average levels of diagnoses in the capital cities of Melbourne and Sydney. The northern Queensland and Northern Territory SA2s with higher than average rates are still visible but use less map space on the hexagon tile map. There are still many light blue areas in both displays. 
-->

## Visual Inference

In order to assess the effectiveness of the hexagon tile map, the lineup protocol [@GIIV,@BCHLLSW09] from visual inference procedures is employed. The procedures for doing a power comparison of competing plot designed, outlined in @GTPCCD, are followed. The approach mirrors classical statistical inference. 

<!--We suggest the hexagon tile map as an alternative display. To test the effectiveness of these displays, there is a formal testing framework. Visual inference considers the communication of data or data summaries through visualizations. It considers these visualizations to be visual statistics, with observable features.
-->

In classical statistical inference hypothesis testing is conducted by comparing the value of a test statistic on a standard reference distribution, computed assuming the null hypothesis is true. If the value is extreme,  the null hypothesis is rejected, because the test statistic value is unlikely to have been so extreme if it was true. In the lineup protocol, the plot plays the role of the test statistic, and the data plot is embedded in a field of null plots. Defining the plot using a grammar of graphics [@ggplot2] makes it a functional mapping of the variables and thus, it can be considered to be a statistic. With the same data, two different plots can be considered to be competing statistics, one possibly a more powerful statistic than the other. 

To do hypothesis testing with the lineup protocol requires human evaluation. The human judge is required to identify the most different plot among the field of plots. If this corresponds to the data plot -- the test statistic -- the null hypothesis is rejected. It means that the data plot is extreme relative to the reference distribution of null plots. 

The null hypothesis is explicitly provided by the grammatical plot description. For example, if a histogram is the plot type being used, the null might be that the underlying distribution of the data is a Gaussian. Null data would be generated by simulating from a normal model, with the same mean and standard deviation as the data. In practice, the null hypothesis used is generic, such as *there is NO structure or a pattern in the plot*, and contrasted to an alternative that there is structure. 

The chance that an observer picks the data plot out of a lineup of size $m$ plots accidentally, if the null hypothesis is true is $1/m$. With $K$ observers, the probability of $k$ randomly choosing the data plot, roughly follows a binomial distribution with $p=1/m$. Fig. \ref{fig:lineup} shows a lineup of the hexagon tile map, of size $m=12$. Plot 3 is the data plot, and the remaining 11 are plots of null data. 

```{r lineup, fig.cap = "This lineup of twelve hexagon tile map displays contains one map with a real population related structure. The rest are null plots that contain spatial correlation between neighbours.", fig.height = 8, fig.width = 8}
ggdraw() +
  draw_plot(rasterGrob(readPNG("figures/aus_cities_3_geo.png")))
```

In order to determine the effectiveness of a type of display, this probability is less relevant than the overall proportion of observers who pick the data plot, $k/K$. The power of the test statistic (data plot) is provided by this proportion. Power in a statistical sense is the ability of the statistic to *produce a rejection* of the null hypothesis, if it is indeed *not true*. With the same data plotted using two different displays, the display with the highest proportion of people who choose the data plot would be considered to be the most powerful statistic. 

<!-- The lineup protocol was formalized as a test for the effectiveness of a visualization. Rather than classical statistical tests, it tasks human participants with finding the display that shows structure between the variables. If they do not detect the structure, then it is not significantly different from the null data.


Visual inference uses the general hypothesis structure: 

 - Null hypothesis: There is \bf{no} relationship between the variables
 - Alternative hypothesis: There is a relationship between the variables

To test these hypotheses, the line up protocol randomly places a "guilty" data visualization in a lineup of "innocents". Where the guilty data set contains structure, and the innocents data sets do not.
In a grid of visualizations, an observer is asked to pick the display that is most different, if they select the data set containing structure, they have identified the guilty hidden within the  innocents [@GTPCCD].
The participants identify the guilty data as different from the innocent data with probability $1/m$, where $m$ is the number of null plots plus 1 to account for the guilty data set. When the guilty data set is chosen we reject the null hypothesis that it was innocent is rejected with a $1/m$ chance or type I error of being wrong.

The lineup protocol can be used in a variety of testing scenarios for visualizations. @GIIV suggest the choropleth map for testing spatial structure in a data set. The lineup protocol allows for flexibility in the definition of null data. In the case of spatial visualizations, it is likely that neighbors are related. To account for this, null data sets can be generated by sampling from a known model that accounts for spatial covariance. 

To contrast the effectiveness of two displays, we can produce null data plots with a hidden real data plot using the different plot designs.
We can compare the time taken by participants to evaluate the same data in the different displays. We can also contrast the accuracy achieved by participants in detecting the real data plot.
-->

Methodology
============

This study aims to answer two key questions around the presentation of spatial distributions:

1. Are spatial disease trends that impact highly populated small areas detected with higher accuracy, when viewed in a hexagon tile map?
2. Are people faster in detecting spatial disease trends that impact highly populated small areas when using a hexagon tile map?

Additional considerations when completing this experimental task included the difficulty experienced by participants and the certainty they had in their decision.

Australia is used for the study, with Statistical Area 2 (SA2) (XXX REF) as the geographic units. The results should apply broadly to any other geographic area of interest. 

<!--
The mean of the detection rate for choropleth map, denoted as $\mu_C$, and the hexagon tile maps, $\mu_H$ will be contrasted. This leads to the following one sided hypothesis:

$H_0$ : $\mu_H$ = $\mu_C$
$H_a$ : $\mu_H$ > $\mu_C$

The detection rate $\hat\pi$ is calculated as the amount of people that made the choice of plot that contained the real data, out of the participants who saw data plot in the lineup of the null data plots.
-->

## Experimental factors

<!-- The variables changed between groups were the type of plot shown and the trend model.-->

The primary factor in the experiment is the plot type. The secondary factor is a trend model. Three trend models were developed, one mirroring a large spatial trend for which the choropleth would be expected to do well, and two with differing level of inner city hot spots. These latter two reflect the structure seen in the liver cancer data (Fig. \ref{fig:liver}). This produces six treatment levels:

  - Map type: *Choropleth, Hexagon tile*
  - Trend: *South-East to North-West; Locations in three population centres; Locations in multiple population centres, *

Data is generated for each of the trend models, with four replicates, and each displayed both as a choropleth and as a hexagon tile map, which yields 12 data sets, and 24 data plots. This set of displays is divided in half, providing two sets of 12 displays, Group A and Group B. Participants were randomly allocated to Group A or B. Participants saw a data set only once, with as a choropleth or as a hexagon tile map. Table \ref{fig:exp-design} summarises the design. (XXX Maybe replace this table with your diagram illustrating the design??)


```{r exp-design, results = "asis", fig.width = 8, fig.height = 8}
ggdraw(xlim = c(0,1), ylim = c(0,1)) + draw_plot(rasterGrob(png::readPNG("figures/experiment_design.png")))
```

## Generating null data

Null data needs to be data with no (interesting) structure. In most scenarios, permutation is the main approach for generating null plots. It is used to break association between variables, while maintaining marginal distributions. This is too simple for spatial data. In spatial data, a key feature is smoothness over the landscape corresponding spatial dependence. To do something simple, like permute the values relative to the geographic location would produce null plots which are too chaotic, and the data plot will be recognisable for its smoothness rather than any structure of interest. 

For spatial data, null data is stationary data, where the mean, variance and spatial dependence are constant over the geographic units. Stationary data is specified by a variogram model (XXX REF). Simulating from a variogram model, where the spatial dependence is specified, generates the stationary spatial data used for the null plots.  The parameters for the Gaussian model were sill=1, range=0.3 with the variance generated by a standard normal distribution. 

The R package `gstat` [@gstat] was used to simulate 144 null sets, 12 data sets for each plot in a lineup, and 12 sets for 12 lineups.

```{r eval=FALSE, echo=FALSE}
var.g.dummy <- gstat(formula = z ~ 1, 
                     locations = ~ longitude + latitude, 
                     dummy = T, beta = 1, model = vgm(psill = 1, model = "Gau", range = 0.3),
                     nmax = 12)
```

<!-- Spatially dependent data sets were simulated using the variogram model on the centroids of each geographic unit. Twelve simulations from the variogram model were created for each of the twelve lineups.


The underlying spatial correlation model was created to provide spatial autocorrelation between neighboring areas using the longitude and latitude values for the Statistical Areas.

$$z = 1$$
$$locations = longitude + latitude$$
-->

## Generating lineups

XXX Fill in this structure - see commented text

<!-- Discuss simulating the trend data -->

<!-- Applying smoothing -->

<!-- Scaling data within a lineup -->

<!-- Location of data plot -->

The null model imposed suggests that neighbors are related. The randomness imposed was smoothed to mirror the practice taken by the Australian Cancer Atlas. This smoothing allowed neighbors to be related to each other and show distributions similar to the Liver cancer distribution shown in \ref{fig:liver}. 

Spatially dependent data sets were simulated using the variogram model on the centroids of each geographic unit. Twelve simulations from the variogram model were created for each of the twelve lineups.

In these 12 sets of data, each of the 144 maps were smoothed several times to replicate the spatial autocorrelation seen in cancer data sets presented in the Australian Cancer Atlas.

For each of the 144 individual maps, the values attributed to each geographic area are rescaled to show a similar color scale from deep blue to dark red within each map.

A random location was selected for each set of lineup data.
In this location, a trend model was overlaid on the null set of spatially correlated data.
Each set of lineup data was used to produce a choropleth maps and hexagon tile maps. These matched pairs were split between Group A and Group B.

For each lineup these statistics were calculated:

 - accuracy: the proportion of subjects who detected the data plot
 - efficiency: average time taken to respond

## Web app collecting responses

XXX Describe the details of the taipan app used to collect responses

Data was collected through a web application containing the online survey.
Each participant used the internet to access the survey.
The data collection took place using a secure link between the survey web application and the googlesheet used to store results. The application would first connect to the googlesheet using the googlesheets [@sheets] R package, and interacted again at the completion of the survey by adding the participant's responses to the 12 displays as 12 rows of data in the googlesheet.

## Participants

Participants were recruited for to evaluate lineups from the Figure Eight crowd sourcing platform [@figeight].
The lineup protocol expects that the participants are uninvolved judges with no prior knowledge of the data, to avoid inadvertently affecting results. Potential participants needed to have achieved level 2 or level 3 from prior work. The participants were chosen as they have experience with survey tasks that involve evaluating images XXX Did we really choose participants this way??). All participants were at least 18 years old.

Participants were allocated to either group A or group B when they proceeded to the survey web application, which was hosted externally to the Figure Eight website. There were 95 participants involved in the study. All participants read introductory materials, and was trained using three test displays, to orient them to the evaluation task. All participants who completed the task were compensated $USD5 for their time, via the Figure Eight payment system.

XXX Sentence on pilot study which allowed us to estimate the effect size, and thus decide on number of participants to collect responses from

## Demographic data collection

The participant answered demographic questions and provided consent before evaluating the lineups.

Demographics were collected regarding the study participants:

 - Gender (female / male / other),
 - Degree education level achieved (high school / bachelors / masters / doctorate / other),
 - Age range (18-24 / 25-34 / 35-44 / 45-54 / 55+ / other)
 - Lived at least for one year in Australia (Yes / No )

Participants then moved to the evaluation phase.
The set of images differed for Group A and Group B.
After being allocated to a group, each individual was shown the 12 displays in randomised order.

Three questions were asked regarding each display:

 - Plot choice
 - Reason
 - Difficulty

After completing the 12 evaluations, the participants were asked to submit their responses.


<!--
## Experimental design


The choropleth map was used as the comparative visualization for presenting the lineups [@VVSIALM] as this is the common display for spatial cancer data.
Geographic distributions usually have some degree of spatial autocorrelation between neighbors.
This feature was incorporated in all maps shown in the lineup displays, the map that contained the trend feature shown in only one set of data was also affected by spatial autocorrelation.
A line up protocol was implemented to arrange $N$ maps in each display. 
A reasonable amount of null plots $N-1$ in the lineup was chosen to ensure the real data map was well hidden. A reasonable number of plots to show in each lineup, $N = 12$ was chosen to not overwhelm participants due to the detailed choropleth maps of Australian SA3 areas.

The hypotheses for each lineup are
$H_0$ : All plots look the same
$H_a$ : One plot looks different to the other plots 


The same data was visualized on a choropleth map, and on a hexagon tile map, however in this study the participants only saw one of these two displays. The accuracy and times taken were aggregated for Groups and.
Comparing the results of participants who see the choropleth to those who see a hexagon tile map will show that population related distributions are spotted more frequently in a hexagon tile map display.


Let $n$ be the number of independent observers and $x_i$ the
number of observers who picked plot $i$, $i = \{1,...,m\}$

Then $x_i, x_2, ..., x_m$ follows a multinomial
distribution$Mult_{\pi_1, \pi_2, ...., \pi_m}(x_i, x_2, ..., x_m)$ with 
$\sum_i \pi_i = 1$, where $\pi_i$ is the probability that plot $i$ is picked by an
observer, which we can estimate as $\hat{\pi}_i = x_i/n$.
The researchers compared the length of time taken, and the accuracy of the participants choices.
The power of a lineup can therefore be estimated as the ratio of correct
identifications $x$ out of $n$ viewings.

## The methods of data analysis used
The data analysis methods used in order to analyse and collate the results included downloading the survey submissions and opening them into the analysis software R [@RCore].
   
For each of the 12 lineup displays the researchers calculated:

 - accuracy: the proportion of subjects who detected the data plot
 - efficiency: average time taken to respond
-->

## Analysis

### Descriptive statistics

Mean, standard deviations for lineups, by participant, ...

Side-by-side dot plots were made of accuracy (efficiency) against map type, faceted by trend model type.

Similar plots were made of the feedback and demographic variables - reason for choice, reported difficulty, gender, age, education, having lived in Australia - against the design variables.

Plots will be made in R (R Core Team 2019), with the ggplot2 package (Wickham 2016).   
   
   
### Modelling

The likelihood of detecting the data plot in the lineup can be modelled using a linear mixed effects model. 
The R [@RCore] `glmer()` function in the [@lme4] package implements generalised linear mixed effect models. The model used includes the two main effects map type and trend model, which gives the fixed effects model to be:


$$\widehat{y_{ij}} = \mu + \tau_i + \delta_j + (\tau\delta)_{ij} + \epsilon_{i,j}, ~~~ i=1,2; ~j=1,2,3$$

where $y_{ij} = 0, 1$ whether the subject detected the data plot, $\mu$ is the overall mean, $\tau_i, i=1,2$ is the map type effect, $\delta_j$ is the trend model effect. We are allowing for an interaction between map type and trend model. Because the response is binary, a logistic model is used. This model can account for each individual participants’ abilities as it includes a subject-specific random intercept. As each participant provides results from 12 lineups.

The model specifies a logistic link, this means the predicted values from the `glmer` model should be back-transformed to fit between 0 and 1. They are transformed with the link specified below:


$$\mu = \frac{e^{\eta}}{1 + e^{\eta}}$$ \label{eq:transform}
$$\eta = f(\tau_i,\delta_j)$$


The feedback and demographic variables will possibly be incorporated as covariates.

Computation will be done using R [@RCore], with the `lme4` package [@lme4].


### Data cleaning

<!-- A pilot study was conducted to determine whether the lineups were appropriate. All lineups in the pilot study were deemed viable as at least one participant detected the real data plot in each lineup.
The demographics of the participants showed a skew towards male participants. 
The randomness of the group allocation also resulted in more participants being allocated to Group B. Due to the allocation of lineup displays the participants all saw six Choropleth displays and six Hexagon Tile Map displays.
-->

Results
============

<!--
The results section explores the [Accuracy] and contrasts the detection rate of the real data plot in the lineups.
The second section considers the [Speed] participants had when submitting their responses to the three questions asked regarding each display.
The [Certainty] participants felt about their decisions and the [Reason] for their choice in plot is also discussed.
-->

Responses from 95 participants were collected. Three participants provided no answers for any task, and their data was removed. Set A was evaluated by 42 participants, and 53 evaluated set B. This resulted in 1104 evaluations, corresponding to 92 subjects, each evaluating 12 lineups, that are analysed on accuracy and speed. The certainty and reasons of subjects in their answers is also examined. 

## Participant demographics

Of the 92 participants, 67 were male, and 25 female. Most participants (56) had a Bachelors degree, 13 had a Masters degree, and the remaining 23 had high school diplomas.

## Accuracy

Fig. \ref{fig:detect-compare} displays the average detection rates for the two types of plot separately for each trend model. Each trend model was tested using four repetitions, evaluations on the same data set were seen as either choropleths or hexagon tile maps by each group as specified in Table. \ref{fig:exp-design}; the detection rates for each display are connected by a line segment. The Three Cities and All Cities trend models shown in the hexagon tile map allowed viewers to detect the data plot substantially more often than the choropleth counterparts. One replicate for the All Cities group, had similar detection rates for both plot type. Surprisingly, participants could also detect the gradual spatial trend in the NW-SE group from the hexagon tile map. We expected that the choropleth map would be superior for the type of spatial pattern, but the data suggests the hexagon tile map performs slightly better, or equally as well.

```{r detect-compare, fig.cap = "The detection rates achieved by participants are contrasted when viewing the four replicates of the three trend models. Each point shows the probability of detection for the lineup display, the facets separate the trend models hidden in the lineup. The points for the same data set shown in a choroleth or haxgon tile map display are linked to show the difference in the detection rate.", fig.height=4}
# Detectability rate for each lineup (image)
d_smry <- d %>% group_by(trend, type, replicate) %>%
  # pdetect measures the aggregated accuracy of the choices
  summarise(pdetect = length(detect[detect == 1])/length(detect)) %>%
  ungroup()

# Numerical summary
diffs <- d_smry %>% spread(type, pdetect) %>%
  mutate(dif = `Hex.` - `Choro.`)

# Plot summary
ggplot(d_smry, aes(x=type, y=pdetect, color = trend)) +
  geom_point(size = 2) +  
  geom_line(size = 1, aes(group = replicate)) +  
  facet_wrap(~trend) +
  scale_color_manual(values = trend_colors) +
  xlab("Type of areas visualized") +
  ylab("Detection rate") + 
  ylim(0,1) +
  guides(color = FALSE)
```


Table. \ref{tab:desc-stats} shows the means and standard deviations of the detection rate for each type of plot and each trend model. This also gives the standard errors, the smallest standard deviation for all sets of replicates is the Three Cities trend model shown in a Choropleth display. This group of displays had a very small detection rate of 0.04. The mean detection rate for the Three Cities trend model shown as choropleth map lineups was also the smallest at 0.39.
The North-West to South-East (NW-SE) trend model unexpectedly had a higher mean detection rate for the hexagon tile map displays, but the difference in the means of detection rate was only 0.10.


```{r desc-stats, results = "asis"}
types <- c("Choro.", "", "Hex.", "")

d %>% group_by(type, trend) %>%
  summarise(m = round(mean(detect),2),
    std.dev = round(sd(detect),2)) %>% 
  mutate(std.dev = ifelse(std.dev == 0.5, "(0.50)", paste0("(", std.dev, ")"))) %>% 
  gather(stat, value, m, std.dev) %>% 
  pivot_wider(names_from = "trend", values_from = "value") %>% 
  arrange(type) %>% ungroup() %>% 
  mutate(Type = types) %>% select(Type, `NW-SE`, `Three Cities`, `All Cities`) %>%
  knitr::kable(., format = "latex", align = "lccc", booktabs = TRUE, 
    linesep = c("", "\\addlinespace"),
    caption = "The rate of detection for each trend model has been calculate for the choropleth and hexagon tile map displays. The associated standard errors are also included.")
```


Table. \ref{tab:detect-glmer1} presents a summary of the generalised linear mixed effects model, testing the effect of plot type and trend model on the detection rate. The results support the summary from Fig. \ref{fig:detect-compare} and all terms are statistically significant despite the large standard deviations observed in Table. \ref{tab:desc-stats}. Overall, the hexagon tile map performs marginally better than the choropleth for all trend models, which is a pleasant surprise. Allowing for the interaction effect, the difference in detection rate decreases for population related displays for a choropleth map lineup, but increases for a hexagon tile map display.
The log odds of detection show in Table. \ref{tab:detect-glmer1} can be back transformed after taking the sum of all terms for the trend and type of display that are of interest.
For the NW-SE distribution, the predicted detection rate for the hexagon tile map display increases the predicted probability of detection to `r round(exp(0.461+0.0220)/(1+exp(0.461+0.0220)), 2)` from `r round(exp(0.0220)/(1+exp(0.0220)),2)` for choropleths, this is almost exactly the difference seen in the table of means and is significant only at the 0.05 level. 

When a choropleth map display is used, the predicted detection rate for the Three Cities trend, `r round(exp(0.0220-3.43)/(1+exp(0.0220-3.43)),2)`; this is extremely low, especially compared to the NW-SE trend of `r round(exp(0.0220)/(1+exp(0.0220)),2)`.
When the All Cities trend is presented in a choropleth display the predicted probability of detection is `r round(exp(0.0220-1.35)/(1+exp(0.0220-1.35)),2)`.
The hexagon tile map has a substantially high detection rate for the display of a Three Cities trend `r round(exp(0.461+0.0220-3.43+2.46)/(1+exp(0.461+0.0220-3.43+2.46)),2)` and All Cities trend `r round(exp(0.461+0.0220-1.35+1.17)/(1+exp(0.461+0.0220-1.35+1.17)),2)`.



```{r detect-glmer1, results="asis"}
# Mixed effects models
glmer1 <- glmer(detect ~ type*trend + (1|contributor), 
              family = binomial, data = d)

glmer_terms <- c("Intercept", "Hex.", "Three Cities", "All Cities", "Hex:Three Cities", "Hex:All Cities")

detection_rates <- tidy(glmer1) %>%
  mutate(detection_rates = round(exp(estimate)/(1+exp(estimate)),2)) %>%
  select(term, estimate, detection_rates) %>% pull(detection_rates)

tidy(glmer1) %>%
  mutate_at(.vars = c("estimate", "std.error"), round, 2) %>% 
  mutate(p.value = round(p.value, digits=2)) %>% 
  rowwise() %>%
  mutate(sig = case_when(
  p.value <= 0.001 ~ "$^{***}$",
  p.value <= 0.01 ~  "$^{**}$",
  p.value <= 0.05 ~  "$^{*}$",
  p.value <= 0.01 ~  "$^{.}$",
  TRUE ~ "$^{ }$")) %>%
  ungroup() %>% 
  filter(!is.na(std.error)) %>%
  mutate(term = glmer_terms) %>% 
  select(Term = term, 
    Est. = estimate, 
    Sig. = sig,
    `Std. Error` = std.error, 
    `P val` = p.value) %>% 
  knitr::kable(format = "latex", escape = FALSE, align= "rrlrr", 
    booktabs = T, linesep = c("", "\\addlinespace", "", "\\addlinespace", ""), 
    caption = "The model output for the generalised linear mixed effect model for detection rate. This model considers the type of display, the trend model hidden in the data plot, and accounts for contributor performance.")
```


## Speed

Fig. \ref{fig:beeswarm} shows horizontally jittered dot plots to contrast the time taken by participants to evaluate each lineup when viewing each type of display. The time are also separated by trend model and whether the data plot was detected or not detected. The time taken to complete an evaluation ranged from milliseconds to 60 seconds. The average time taken for type of display is shown as a large colored dot on each plot. when considering the heights of the green and orange dots, there is little difference in the average time taken to read a choropleth or hexagon tile map. Comparing the same colored dot across each trend model row, there is a slight increase in the time taken to correctly detected the data plot in the hexagon tile map lineup, but little difference in evaluation time for the choropleth display. However, there were substantially less correct detections for choropleth lineups for the Three cities and All Cities trends.


```{r beeswarm, fig.cap = "The distribution of the time taken (seconds) to submit a response for each combination of trend, whether the data plot was detected, and type of display, shown using horizontally jittered dotplots. The colored point indicates average time taken for each plot type. Although some participants take just a few seconds per evaluation, and some take as much as mcuh as 60 seconds, but there is very little difference in time taken between plot types.", fig.height=8}
# Di playing
s <- d %>% group_by(type, trend, detect_f) %>%
  summarise(m=median(time_taken), 
            q1=quantile(time_taken, 0.25), 
            q3=quantile(time_taken, 0.75))
library(ggbeeswarm)
ggplot() + 
  geom_quasirandom(data=d, aes(x=type, y=time_taken), alpha=0.9) + 
  #geom_hline(data=s, aes(yintercept=m, color=type)) +
  geom_point(data=s, aes(x=type, y=m, color=type), size=5, alpha=0.7) +
  #geom_errorbar(data=s, aes(x=type, ymin=q1, ymax=q3, color=type), width=0.3, size=2) +
  scale_color_brewer("", palette = "Dark2") +
  facet_grid(trend~detect_f) + 
  ylab("Time taken (seconds)") + xlab("") +
  theme(legend.position="bottom")
```


## Certainty

Participants provided their level of certainty regarding their choice using a five point scale. 
Unlike the accuracy and speed of responses that were derived during the data processing phase, this was a subjective
assessment by the participant prompted by the question: ‘How certain are you about your choice?’.
Figure. \ref{certainty} shows the amount of times participants provided each level of certainty. This was separated for each combination of trend models and display type, and colored depending on whether a participant correctly detected the data plot in the lineup.
Participants chose 4 or 5 often when viewing the population related trends in the choropelth display, even though they were often incorrect when viewing an All Cities trend and overwhelmingly incorrect for the Three Cities trend. This shows overconfidence in their detection ability when using a choropleth map display. Participants were less likely to be certain when their choice was incorrect and they were viewing a hexagon tile map.
For each trend model, participants were more likely to doubt their choice and choose 1 or 2 in the hexagon tile map displays, even though many had made the correct choice. 


```{r certainty, fig.cap = "The amount of times each level of certainty was chosen by participants when viewing hexagon tile map or choropleth displays. Participants were more likely to choose a high certainty when considering a Choropleth map. The mid value of 3 was the default certainty, it was chosen most for the Hexagon tile map displays.", fig.height=5}
d <- d %>% 
  mutate(certainty = as_factor(certainty)) %>% 
  mutate(replicate_f = as_factor(replicate)) 

ggplot(d %>% rename(Detected = detect_f), 
   aes(x = certainty, fill = Detected)) +  
  scale_fill_manual(values = detect_colors) +
  geom_bar() + facet_grid(type ~ trend) +
  theme(legend.position = "bottom") + xlab("Level of certainty")
```


## Reason

Participants were asked why they had made their plot choice and were able to select from a set of suggested reasons. 
"Color trend across the areas" was the most common selection for NW-SE trend displays.

The reasons used when viewing choropleth displays varied more than the hexagon tile map reasons.
The hexagon tile map displays resulted in "Clusters of color" as the most common choice made by participants.

The choice "None of these reasons" was used as the default value to minimise noise from participants who did not select a response. 

```{r reason, results = "asis"}
# Qualitative analysis of reason
d %>% 
  mutate(reason = ifelse(reason =="0.0", "no reason", reason)) %>% 
  mutate(Detected = ifelse(detect_f == "Detected? Yes", "Yes", "No"),
    Trend = trend) %>% 
  group_by(Trend, Detected, type) %>% 
  count(reason) %>% 
  mutate(prop = round(n/sum(n), 2), r_prop = paste0(reason, ":", prop)) %>% 
  top_n(1, n) %>% summarise(reasons = paste(reason, collapse=", ")) %>% 
  pivot_wider(names_from = c("type"), values_from = c("reasons")) %>% 
  ungroup() %>% 
  knitr::kable(., format = "latex", booktabs = TRUE,
    linesep = c("", "\\addlinespace", "", "\\addlinespace",""),
    caption = "The amount of participants that selected each reason for their choice of plot when looking at each trend model shown in Choropleth and Hexagon Tile maps. The facets show whether or not the choice was correct.") %>% 
  #kable_styling(bootstrap_options = "hold_position") %>% 
  collapse_rows(., columns = 1)
```



Discussion
============

The intention of this study was to contrast the use of the choropleth map and the hexagon tile map.
The visual inference lineup protocol was employed to contrast the effectiveness of the displays.
The results have shown that overall the use of the hexagon tile map display allows participants to find the data plot in the lineup more often.
Using the visual inference protocol this result can be extended to show that it is a valid alternative display to communicate spatial distributions of population related data.

We expected that the choropleth map would be superior for communicating the spatial pattern of geographic distributions. The data suggest that the participants perform slightly better or equally as well for each replicate in each trend model across the two displays. Table II shows that the difference in the mean detection rate for the two trend models was 0.10.

The differences seen in the compare detect plot and Table. \ref{tab:desc-stats} are reflected in the model results. Surprisingly the difference scene for the geographic distribution was significant at the 0.05 level.
It also showed that the hexagon tile map display performs marginally better than the choropleth for all trend models. Unexpectedly the detection rate suffers when using a choropleth map to display population related distributions.

While the significance of the difference in detection was the key focus of this experiment, the secondary focus was the time taken by participants. it was expected that the disciplines may take longer to consider the hexagon tile map distribution but would be able to detect the data plot in the lineup.
The bimodal distributions seen in the Fig. \ref{fig:beeswarm} display showed very little difference in the mean evaluation times. As the ranges of all of the distributions approached 60 seconds it cannot be said that the participants' took longer to evaluate the hexagon tile map displays. 

The responses to the questions asked of participants included the reason for their choice and the certainty around their choice.
Fig. \ref{fig:reason} shows high levels of certainty of 4 and 5 were chosen by participants when looking at the population distributions in a choropleth map display show that they were over confident when attempting to find the real data plot in the choropleth map displays. Participants performed better on the NW-SE distribution shown in the choropleth display and were reasonably confident about their decisions.
The high levels of the mid range value of 3 could indicate that the participant did not want to provide a response, as this was the default value. Those who chose level 4 or 5 were equally likely to be correct for the three cities lineups, but more likely to be correct than incorrect for the other two trend models.


The color scaling applied in Three cities and All cities displays resulted in the rural areas of the real data plot appearing more blue or yellow than the other plots in the lineups.
Due to the consistent coloring of rural areas in a choropleth display, the choice "All areas have similar colors" was most common reason for a participants choice. The All Cities displays colored the inner-city areas of all capital cities more red, this was observable to participants and explains the equal choice of the city clusters or rural color consistency. 
Choosing "Clusters of colour" was expected when participants viewed the Hexagon tile map display of the All Cities and Three Cities distributions. It was unexpected that it was also the most common reason for the NW-SE hexagon tile map displays. 
Due to the spatial covariance introduced in the smoothing, groups of similarly colored hexagons were present in all of the hexagon tile map displays. All Cities and Three Cities distributions of real data trends had distinctly different patterns or red inner-city areas, while some of the plots in each lineup may have shared similar features.


Conclusion
============

The choropleth map display and the tessellated hexagon tile map have been contrasted using the lineup protocol. The hexagon tile map was significantly more effective for spotting a real population related data trend model hidden in a lineup.

The hexagon tile map display should be considered as an alternative visualization method when communicating distributions that relate to the population across a set of geographic units. As an additional display to the familiar choropleth map, Cancer Atlas products may benefit from the opportunity to allow exploration via an alternative display. The spatial distributions used to test these displays were inspired by the real spatially smoothed estimates of the cancer burden on Australian communities. However, this technique may be extended to other population related distributions, such as other diseases.

The increasing population densities of capital cities despite large land area exacerbates the difference in the smallest and largest communities.
The population density structure of Australia can be considered similar to that of Canada, New Zealand and many others. Therefore, this display is not only relevant to Australia, but all nations or population distributions that experience densely populated cities separated by vast rural expanses.



Acknowledgements
==============

The authors would like to thank the Australian Cancer Atlas team for discussions regarding alternative spatial visualizations. They would also like to thank Kerrie Mengersen and Dr. Earl Duncan for suggestions and comments.

Ethics approval for the online survey was granted by QUT’s Ethics Committee (Ethics Application Number: 1900000991). All applicants provided informed consent in line with QUT regulations prior to participating in this research. 

All code to generate the data, of the web app that collected participants responses, conduct the analysis, and produce this manuscript are available on XXX. 

The supplementary material contains:

- Additional analysis of the experimental results
- Training materials used for the participants
- 24 lineups used in the experiment
- 12 data sets used to construct the lineups

References {#references .numbered}
==========



<!-- An example of a floating figure using the graphicx package. -->
<!-- Note that \label must occur AFTER (or within) \caption. -->
<!-- For figures, \caption should occur after the \includegraphics. -->
<!-- Note that IEEEtran v1.7 and later has special internal code that -->
<!-- is designed to preserve the operation of \label within \caption -->
<!-- even when the captionsoff option is in effect. However, because -->
<!-- of issues like this, it may be the safest practice to put all your -->
<!-- \label just after \caption rather than within \caption{}. -->

<!-- Reminder: the "draftcls" or "draftclsnofoot", not "draft", class -->
<!-- option should be used if it is desired that the figures are to be -->
<!-- displayed while in draft mode. -->

<!-- \begin{figure}[!t] -->
<!-- \centering -->
<!-- \includegraphics[width=2.5in]{myfigure} -->
<!-- where an .eps filename suffix were assumed under latex,  -->
<!-- and a .pdf suffix were assumed for pdflatex; or what has been declared -->
<!-- via \DeclareGraphicsExtensions. -->
<!-- \caption{Simulation results for the network.} -->
<!-- \label{fig_sim} -->
<!-- \end{figure} -->

<!-- Note that the IEEE typically puts floats only at the top, even when this -->
<!-- results in a large percentage of a column being occupied by floats. -->


<!-- An example of a double column floating figure using two subfigures. -->
<!-- (The subfig.sty package must be loaded for this to work.) -->
<!-- The subfigure \label commands are set within each subfloat command, -->
<!-- and the \label for the overall figure must come after \caption. -->
<!-- \hfil is used as a separator to get equal spacing. -->
<!-- Watch out that the combined width of all the subfigures on a  -->
<!-- line do not exceed the text width or a line break will occur. -->

<!-- \begin{figure*}[!t] -->
<!-- \centering -->
<!-- \subfloat[Case I]{\includegraphics[width=2.5in]{box}% -->
<!-- \label{fig_first_case}} -->
<!-- \hfil -->
<!-- \subfloat[Case II]{\includegraphics[width=2.5in]{box}% -->
<!-- \label{fig_second_case}} -->
<!-- \caption{Simulation results for the network.} -->
<!-- \label{fig_sim} -->
<!-- \end{figure*} -->

<!-- Note that often IEEE papers with subfigures do not employ subfigure -->
<!-- captions (using the optional argument to \subfloat[]), but instead will -->
<!-- reference/describe all of them (a), (b), etc., within the main caption. -->
<!-- Be aware that for subfig.sty to generate the (a), (b), etc., subfigure -->
<!-- labels, the optional argument to \subfloat must be present. If a -->
<!-- subcaption is not desired, just leave its contents blank, -->
<!-- e.g., \subfloat[]. -->

\newpage

Acknowledgment
==============

The authors would like to thank the Australian Cancer Atlas team for discussions regarding alternative spatial visualizations, and Professor Kerrie Mengersen and Dr Earl Duncan for regular meetings filled with suggestions and comments. Mitchell O'Hara-Wild was a co-developer of the taipan [@taipan] R package for image tagging, used as the base for the web app constructed to collect participant evaluations of lineups. We are thankful for the NUMBATs (Non-Uniform Monash Business Analytics Team) for participating in the pilot study that helped to assess the experimental design and determine an appropriate sample size for the study. 

The source code to produce this document can be found on [GitHub](https://github.com/srkobakian/experiment/tree/master/paper).
Supplementary materials have been included to discuss the survey procedures and the lineups that were used.
The full set of images can also be found at: https://github.com/srkobakian/experiment/tree/master/figures/final. 

The analysis of the work was completed in R [@RCore] with the use of the following packages:

 - For document creation: rmarkdown [@rmarkdown], rticles [@rticles], knitr [@knitr].
 - For lineup creation and data analysis: tidyverse [@tidyverse], nullabor [@nullabor], ggthemes [@ggthemes], RColorBrewer [@RColorBrewer].
 - For image displays: cowplot [@cowplot], png [@png], grid [@grid].
 - For modelling and presentation of models: gstat [@gstat], lme4 [@lme4], kableExtra [@kableExtra].

Ethics approval for the online survey was granted by QUT’s Ethics Committee (Ethics Application Number: 1900000991). All applicants provided informed consent in line with QUT regulations prior to participating in this research. 


\newpage
References {.numbered}
==========



<!-- An example of a floating figure using the graphicx package. -->
<!-- Note that \label must occur AFTER (or within) \caption. -->
<!-- For figures, \caption should occur after the \includegraphics. -->
<!-- Note that IEEEtran v1.7 and later has special internal code that -->
<!-- is designed to preserve the operation of \label within \caption -->
<!-- even when the captionsoff option is in effect. However, because -->
<!-- of issues like this, it may be the safest practice to put all your -->
<!-- \label just after \caption rather than within \caption{}. -->

<!-- Reminder: the "draftcls" or "draftclsnofoot", not "draft", class -->
<!-- option should be used if it is desired that the figures are to be -->
<!-- displayed while in draft mode. -->

<!-- \begin{figure}[!t] -->
<!-- \centering -->
<!-- \includegraphics[width=2.5in]{myfigure} -->
<!-- where an .eps filename suffix were assumed under latex,  -->
<!-- and a .pdf suffix were assumed for pdflatex; or what has been declared -->
<!-- via \DeclareGraphicsExtensions. -->
<!-- \caption{Simulation results for the network.} -->
<!-- \label{fig_sim} -->
<!-- \end{figure} -->

<!-- Note that the IEEE typically puts floats only at the top, even when this -->
<!-- results in a large percentage of a column being occupied by floats. -->


<!-- An example of a double column floating figure using two subfigures. -->
<!-- (The subfig.sty package must be loaded for this to work.) -->
<!-- The subfigure \label commands are set within each subfloat command, -->
<!-- and the \label for the overall figure must come after \caption. -->
<!-- \hfil is used as a separator to get equal spacing. -->
<!-- Watch out that the combined width of all the subfigures on a  -->
<!-- line do not exceed the text width or a line break will occur. -->

<!-- \begin{figure*}[!t] -->
<!-- \centering -->
<!-- \subfloat[Case I]{\includegraphics[width=2.5in]{box}% -->
<!-- \label{fig_first_case}} -->
<!-- \hfil -->
<!-- \subfloat[Case II]{\includegraphics[width=2.5in]{box}% -->
<!-- \label{fig_second_case}} -->
<!-- \caption{Simulation results for the network.} -->
<!-- \label{fig_sim} -->
<!-- \end{figure*} -->

<!-- Note that often IEEE papers with subfigures do not employ subfigure -->
<!-- captions (using the optional argument to \subfloat[]), but instead will -->
<!-- reference/describe all of them (a), (b), etc., within the main caption. -->
<!-- Be aware that for subfig.sty to generate the (a), (b), etc., subfigure -->
<!-- labels, the optional argument to \subfloat must be present. If a -->
<!-- subcaption is not desired, just leave its contents blank, -->
<!-- e.g., \subfloat[]. -->
